---
title: "Bulk clustering analysis"
author: "Alonso Serrano"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(stringr)
conditions <- c("control", "fertilised", "combined")

#specie <- "cenococcum_geophilum"
#specie <- "cortinarius_glaucopus"
specie <- "piloderma_croceum"
speciesName <- gsub("_"," ",str_to_title(specie))

library(data.table)
library(dplyr)
library(here)

# preload packages for depencies
library(ggplot2)
library(RColorBrewer)
library(reshape2)
library(tidyr)
library(tibble)
library(jsonlite)
library(treemap)
library(KEGGREST)
library(reshape2)

source("~/Git/Rtoolbox/src/plotEigenGene.R")
source("~/Git/Rtoolbox/src/getEigengenes.R")
source("~/Git/Rtoolbox/src/plotEnrichedTreemap.R")
source("~/Git/UPSCb-common/src/R/gopher.R")
source("~/scripts/koPathwayDiseaesCleanup/koProcessing.R")

spruceDir <- "/mnt/picea/projects/spruce/vhurry/14_SpruceRoots_Project"
fungiDir <- file.path("/mnt/picea/projects/spruce/vhurry/fungi-flakaliden-2011/fungi-species",specie)

spruceDE <- file.path(spruceDir,"DE")
fungiDE <- file.path(fungiDir,"DE")

checkFile <- function(fileName) {
  if (!file.exists(fileName)) stop(paste("File", fileName, "doesn't exist"))
}
```

# SETUP
## Obtain profile degradation matrix
```{r}
load(here("env", paste0(specie, "_controlDEG.RData")))
```

## Load expression data
```{r loadRdata}
load(file.path(spruceDE, "controlVsdData.RData"))
controlSpruceData <- controlData
load(file.path(spruceDE, "fertilisedVsdData.RData"))
fertilisedSpruceData <- fertilisedData

load(file.path(fungiDE, "controlVsdData.RData"))
controlFungiData <- controlData
load(file.path(fungiDE, "fertilisedVsdData.RData"))
fertilisedFungiData <- fertilisedData
```

## Load network data
```{r}
# load spruce data files
controlSpruceClusterFile <- file.path(spruceDir,"spruce-networks/control/cluster","InfomapClusters.tsv")
checkFile(controlSpruceClusterFile)

# load fungi data files
controlFungiClusterFile <- file.path(fungiDir,"networks/control/results/aggregated","infomapClusters.tsv")
checkFile(controlFungiClusterFile)
```

## Load files
```{r loadFiles}
#name correction, spruce rownames, have an extra dash
rownames(controlSpruceData) <- gsub("-", "", rownames(controlSpruceData))
rownames(fertilisedSpruceData) <- gsub("-", "", rownames(fertilisedSpruceData))

controlSpruceClusters <- read.table(controlSpruceClusterFile, stringsAsFactors=F, header=T, sep='\t', comment.char="")
controlFungiClusters <- read.table(controlFungiClusterFile, stringsAsFactors=F, header=T, sep='\t', comment.char="")

```

## Split the clusters
```{r splitClusters}
spruceClusterNames<- unique(controlSpruceClusters$cluster)
splitSpruceClusters <- lapply(spruceClusterNames, function(cluster){
  theCluster <- controlSpruceClusters[controlSpruceClusters$cluster==cluster,]$gene
  res <- getEigenGenes(controlSpruceData, theCluster)
  res
})
names(splitSpruceClusters) <- spruceClusterNames


fungiClusterNames<- unique(controlFungiClusters$cluster)
splitFungiClusters <- lapply(fungiClusterNames, function(cluster){
  theCluster <- controlFungiClusters[controlFungiClusters$cluster==cluster,]$gene
  res <- getEigenGenes(controlFungiData, theCluster)
  res
})
names(splitFungiClusters) <- fungiClusterNames

```

# Bulk analysis  
4 criteria  

I - Always coordinated: >0.5 in both conditions OR <-0.5 in both conditions  
II - Never coordinated: >-0.5 AND <0.5 in both conditions  
III - Improved coordination: >-0.5 AND <0.5 in ND AND >0.5 OR <-0.5 in NE  
IV - Deteriorated coordination: >0.5 OR <-0.5 in ND AND >-0.5 AND <0.5 in NE  
```{r}
criteria1 <- controlDEG %>% filter((control_score > 0.5 & fertilised_score > 0.5) | (control_score < -0.5 & fertilised_score < -0.5) )
criteria2 <- controlDEG %>% filter((control_score < 0.5 & fertilised_score < 0.5) & (control_score > -0.5 & fertilised_score > -0.5) )
criteria3 <- controlDEG %>% filter(control_score > -0.5 & control_score < 0.5 & (fertilised_score > 0.5 | fertilised_score < -0.5) )
criteria4 <- controlDEG %>% filter((control_score > 0.5 |  control_score < -0.5) & (fertilised_score > -0.5 & fertilised_score < 0.5) )
sum(dim(criteria1)[1] + dim(criteria2)[1] + dim(criteria3)[1] + dim(criteria4)[1])

write.table(criteria1, file=paste0(specie,"_criteria1.tsv"), quote = F, row.names=F)
write.table(criteria2, file=paste0(specie,"_criteria2.tsv"), quote = F, row.names=F)
write.table(criteria3, file=paste0(specie,"_criteria3.tsv"), quote = F, row.names=F)
write.table(criteria4, file=paste0(specie,"_criteria4.tsv"), quote = F, row.names=F)
```

### Clusterize function
```{r}
bulkClusterize <- function(criteria, spruceClusters, fungiClusters) {
  
  res <- list()
  
  res$spruceRes <- c()
  
  None <- sapply(unique(criteria[["Spruce-control"]]), function(x){
  res2 <- if(grepl("i", x) ) {
      spruceClusters[[x]]$negative
    } else {
      spruceClusters[[x]]$positive
    }
  
  res$spruceRes <<- c(res$spruceRes, res2)
  })
  #names(res$spruceRes) <- unique(criteria[["Spruce-control"]])
  #res$spruceRes <- melt(res$spruceRes)
  
  res$fungiRes <-  c() 
    None <- sapply(unique(criteria[["Fungi-control"]]), function(x){
    res2 <- if(grepl("i", x)){
      fungiClusters[[x]]$negative
    } else {
      fungiClusters[[x]]$positive
    }
    res$fungiRes <<- c(res$fungiRes, res2)
  })
    #names(res$fungiRes) <- unique(criteria[["Fungi-control"]])
 # res$fungiRes <- melt(res$fungiRes)

  
  return(res)
}
```

```{r}
megaCluster <- bulkClusterize(criteria1, splitSpruceClusters, splitFungiClusters )
```

# SPRUCE
## Expression profile
```{r}
plotEigengene(controlSpruceData,megaCluster$spruceRes, condition = substr(rownames(controlSpruceData), 4,4), time = as.integer(substr(rownames(controlSpruceData), 2,3)))
```


```{r eval=FALSE,include=FALSE}
load(here("env/spruceMegaClusterEnr.RData"))
```

```{r eval=T, include=T}
spruceEnr <- gopher(megaCluster$spruceRes, background = controlSpruceClusters$gene, url='pabies', task=c("go","kegg", "mapman", "pfam"))
```

```{r eval=FALSE, include=FALSE}
save(spruceEnr, file=here("env/spruceMegaClusterEnr.RData"))
```

## Spruce treemaps
```{r message=FALSE, warning=FALSE}
if(!is.null(spruceEnr$go))
  plotEnrichedTreemap(spruceEnr, "go", "none", title = "GO")
if(!is.null(spruceEnr$kegg)) {
  spruceEnr$kegg <- koTranslate(spruceEnr$kegg)
  plotEnrichedTreemap(spruceEnr, "kegg", "none", title = "KEGG")
}
if(!is.null(spruceEnr$mapman))
  plotEnrichedTreemap(spruceEnr, "mapman", "none", title = "MAPMAN")

if(!is.null(spruceEnr$pfam))  
  plotEnrichedTreemap(spruceEnr, "pfam", "none", title = "PFAM")
```

# FUNGI
## Expression profile
```{r}
plotEigengene(controlFungiData,megaCluster$fungiRes, condition = substr(rownames(controlFungiData), 4,4), time = as.integer(substr(rownames(controlFungiData), 2,3)))
```

```{r}
fungiEnr <- gopher(megaCluster$fungiRes, background = controlFungiClusters$gene, url=paste0('fungi2011'), task=c("cog","ko", "kog", "ko_pathway"))
fungiGOEnr <- gopher(megaCluster$fungiRes, background = controlFungiClusters$gene, url=paste0('fungi2011_',specie), task=c("go"))
```

## Fungi treemaps
```{r  message=FALSE, warning=FALSE}
if(!is.null(fungiEnr$cog))
  plotEnrichedTreemap(fungiEnr, "cog", "none", title = "COG")
if(!is.null(fungiEnr$ko)) {
  fungiEnr$ko <- koTranslate(fungiEnr$ko)
  plotEnrichedTreemap(fungiEnr, "ko", "none", title = "KO")
}

if(!is.null(fungiEnr$ko_pathway)) {
  fungiEnr$ko_pathway <- koPathwayDiseaseCleanup(koTranslate(fungiEnr$ko_pathway))
  plotEnrichedTreemap(fungiEnr, "ko_pathway", "none", title = "KO PATHWAY")
}

if(!is.null(fungiEnr$kog))
  plotEnrichedTreemap(fungiEnr, "kog", "none", title = "KOG")

if(!is.null(fungiGOEnr$go))  
  plotEnrichedTreemap(fungiGOEnr, "go", "none", title = "GO")
```
# Session info
```{r}
sessionInfo()
```