---
title: "Flakaliden 2011 - DE comparatives"
author: "Alonso Serrano"
date: "`r Sys.Date()`"
output:
 html_document:
   toc: true
   number_sections: true
---


# Introduction
This file contains the script and results for the comparatived analysis done to 
the spruce roots and fungi community in Flakaliden 2011 data. 

*201005 Creation

## Prerequisites

# Setup
Load libraries
```{r, message=FALSE}
library(DESeq2)
library(tximport)
library(BiocParallel)
library(plotly)
library(dplyr)
library(here)
library(limma)
register(MulticoreParam(4))
source(here("Rtoolbox/src/utilsDE.r"))
source(here("Rtoolbox/src/plot3Dvector.R"))
source(here("Rtoolbox/src/plotVectorPCA.R"))
source(here("Rtoolbox/src/plotUpAndDown.R"))
```

Set the project folder

```{r}
projectSpruceFolder <- "/mnt/picea/projects/spruce/vhurry/14_SpruceRoots_Project"
deSpruceFolder <- file.path(projectSpruceFolder, "DE")

projectFungiKOFolder <- "/mnt/picea/projects/spruce/vhurry/fungi-flakaliden-2011/"
deFungiKOFolder <- file.path(projectFungiKOFolder, "DE")
```

Auxiliar functions
```{r}
getUpDownGenes <- function(res) {
  res.filter <- filterDE(res)
  upGenes <- rownames(res.filter[res.filter$log2FoldChange > 0,])
  downGenes <- rownames(res.filter[res.filter$log2FoldChange < 0,])
  return(list(up=upGenes, down =downGenes))
}
```

# SPRUCE

Load precalculated DESeq2 results
```{r}
load(file.path(deSpruceFolder, "ddsTreatment.RData"))
load(file.path(deSpruceFolder, "allDEresults.RData"))
spruceTreatmentResults <- results(ddsSpruceTreatment)
spruceControlTimeResults <- control.res
spruceFertilisedTimeResults <- fertilised.res
spruceCombinedResults <- combined.res
```

## Spruce Treatment DE
```{r}
upSpruceTreatment <- getUpDownGenes(spruceTreatmentResults)$up
downSpruceTreatment <- getUpDownGenes(spruceTreatmentResults)$down
```
* Number of genes up regulated due to treatment  `r length(upSpruceTreatment)`
* Number of genes down regulated  due to treatment `r length(downSpruceTreatment)`

## Spruce Control Time DE
```{r}
upDownSpruceTime <- lapply(spruceControlTimeResults, getUpDownGenes)

onceUp <- c()
onceDown <- c()
  
for (i in upDownSpruceTime) {
  onceUp <- union(onceUp, i$up)
  onceDown <- union(onceDown, i$down)
}

upAndDown <- intersect(onceUp, onceDown)
```
* Number of genes up regulated in at least one time point `r length(onceUp)`
* Number of genes down regulated in at least one time point `r length(onceDown)`
* Number of genes only up regulated in at least one time point `r length(setdiff(onceUp, onceDown))`
* Number of genes only down regulated in at least one time point `r length(setdiff(onceDown, onceUp))`

## Spruce Fertilised Time DE
```{r}
upDownSpruceTime <- lapply(spruceFertilisedTimeResults, getUpDownGenes)

onceUp <- c()
onceDown <- c()
  
for (i in upDownSpruceTime) {
  onceUp <- union(onceUp, i$up)
  onceDown <- union(onceDown, i$down)
}

upAndDown <- intersect(onceUp, onceDown)

```
* Number of genes up regulated in at least one time point `r length(onceUp)`
* Number of genes down regulated in at least one time point `r length(onceDown)`
* Number of genes only up regulated in at least one time point `r length(setdiff(onceUp, onceDown))`
* Number of genes only down regulated in at least one time point `r length(setdiff(onceDown, onceUp))`


# FUNGI

Load precalculated DESeq2 results
```{r}
load(file.path(deFungiKOFolder, "ddsTreatment_ko.RData"))
load(file.path(deFungiKOFolder, "allDEresults.RData"))
fungiTreatmentResults <- results(ddsFungiTreatment)
fungiControlTimeResults <- control.res
fungiFertilisedTimeResults <- fertilised.res
fungiCombinedResults <- combined.res
```


## Fungi Treatment DE
```{r}
upFungiTreatment <- getUpDownGenes(fungiTreatmentResults)$up
downFungiTreatment <- getUpDownGenes(fungiTreatmentResults)$down
```

* Number of KOs up regulated due to treatment  `r length(upFungiTreatment)`
* Number of KOs down regulated  due to treatment `r length(downFungiTreatment)`

## Fungi Control Time DE
```{r}
upDownFungiTime <- lapply(fungiControlTimeResults, getUpDownGenes)

onceUp <- c()
onceDown <- c()
  
for (i in upDownFungiTime) {
  onceUp <- union(onceUp, i$up)
  onceDown <- union(onceDown, i$down)
}

upAndDown <- intersect(onceUp, onceDown)

```
* Number of KOs up regulated in at least one time point `r length(onceUp)`
* Number of KOs down regulated in at least one time point `r length(onceDown)`
* Number of KOs only up regulated in at least one time point `r length(setdiff(onceUp, onceDown))`
* Number of KOs only down regulated in at least one time point `r length(setdiff(onceDown, onceUp))`

## Fungi Fertilised Time DE
```{r}
upDownFungiTime <- lapply(fungiFertilisedTimeResults, getUpDownGenes)

onceUp <- c()
onceDown <- c()
  
for (i in upDownFungiTime) {
  onceUp <- union(onceUp, i$up)
  onceDown <- union(onceDown, i$down)
}

upAndDown <- intersect(onceUp, onceDown)

```
* Number of KOs up regulated in at least one time point `r length(onceUp)`
* Number of KOs down regulated in at least one time point `r length(onceDown)`
* Number of KOs only up regulated in at least one time point `r length(setdiff(onceUp, onceDown))`
* Number of KOs only down regulated in at least one time point `r length(setdiff(onceDown, onceUp))`

# LFC magnitude

## Treatment
```{r}
load(file.path(deSpruceFolder, "ddsTreatment.RData"))
load(file.path(deFungiKOFolder, "ddsTreatment_ko.RData"))
# eliminate NAs

resFungiTreatment$FertilisedvsControl$log2FoldChange <-
  ifelse(is.na(resFungiTreatment$FertilisedvsControl$log2FoldChange), 0, resFungiTreatment$FertilisedvsControl$log2FoldChange)

resSpruceTreatment$FertilisedvsControl$log2FoldChange <-
  ifelse(is.na(resSpruceTreatment$FertilisedvsControl$log2FoldChange), 0, resSpruceTreatment$FertilisedvsControl$log2FoldChange)

treatmentList <- list(SpruceFvsC = resSpruceTreatment$FertilisedvsControl$log2FoldChange,
                      FungiFvsC = resFungiTreatment$FertilisedvsControl$log2FoldChange)

boxplot(treatmentList,ylim=c(-5,5), col=c("#00FF00", "#FF0000"), ylab="log2FoldChange" )


```
## Control time
```{r}
# spruceControlTimeList <- list(Spruce20Cvs19C = spruceControlTimeResults[["20Cvs19C"]]$log2FoldChange,
#                               Spruce21Cvs20C = spruceControlTimeResults[["21Cvs20C"]]$log2FoldChange,
#                               Spruce22Cvs21C = spruceControlTimeResults[["22Cvs21C"]]$log2FoldChange,
#                               Spruce23Cvs22C = spruceControlTimeResults[["23Cvs22C"]]$log2FoldChange,
#                               Spruce24Cvs23C = spruceControlTimeResults[["24Cvs23C"]]$log2FoldChange,
#                               Spruce25Cvs24C = spruceControlTimeResults[["25Cvs24C"]]$log2FoldChange,
#                               Spruce26Cvs25C = spruceControlTimeResults[["26Cvs25C"]]$log2FoldChange,
#                               Spruce28Cvs26C = spruceControlTimeResults[["28Cvs26C"]]$log2FoldChange,
#                               Spruce31Cvs28C = spruceControlTimeResults[["31Cvs28C"]]$log2FoldChange,
#                               Spruce32Cvs31C = spruceControlTimeResults[["32Cvs31C"]]$log2FoldChange,
#                               Spruce34Cvs32C = spruceControlTimeResults[["34Cvs32C"]]$log2FoldChange,
#                               Spruce35Cvs34C = spruceControlTimeResults[["35Cvs34C"]]$log2FoldChange,
#                               Spruce36Cvs35C = spruceControlTimeResults[["36Cvs35C"]]$log2FoldChange,
#                               Spruce37Cvs36C = spruceControlTimeResults[["37Cvs36C"]]$log2FoldChange,
#                               Spruce38Cvs37C = spruceControlTimeResults[["38Cvs37C"]]$log2FoldChange,                             
#                               Spruce39Cvs38C = spruceControlTimeResults[["39Cvs38C"]]$log2FoldChange,
#                               Spruce40Cvs39C = spruceControlTimeResults[["40Cvs39C"]]$log2FoldChange,
#                               Spruce41Cvs40C = spruceControlTimeResults[["41Cvs40C"]]$log2FoldChange)
# 
# fungiControlTimeList <- list(Fungi20Cvs19C = fungiControlTimeResults[["20Cvs19C"]]$log2FoldChange,
#                               Fungi21Cvs20C = fungiControlTimeResults[["21Cvs20C"]]$log2FoldChange,
#                               Fungi22Cvs21C = fungiControlTimeResults[["22Cvs21C"]]$log2FoldChange,
#                               Fungi23Cvs22C = fungiControlTimeResults[["23Cvs22C"]]$log2FoldChange,
#                               Fungi24Cvs23C = fungiControlTimeResults[["24Cvs23C"]]$log2FoldChange,
#                               Fungi25Cvs24C = fungiControlTimeResults[["25Cvs24C"]]$log2FoldChange,
#                               Fungi26Cvs25C = fungiControlTimeResults[["26Cvs25C"]]$log2FoldChange,
#                               Fungi28Cvs26C = fungiControlTimeResults[["28Cvs26C"]]$log2FoldChange,
#                               Fungi31Cvs28C = fungiControlTimeResults[["31Cvs28C"]]$log2FoldChange,
#                               Fungi32Cvs31C = fungiControlTimeResults[["32Cvs31C"]]$log2FoldChange,
#                               Fungi34Cvs32C = fungiControlTimeResults[["34Cvs32C"]]$log2FoldChange,
#                               Fungi35Cvs34C = fungiControlTimeResults[["35Cvs34C"]]$log2FoldChange,
#                               Fungi36Cvs35C = fungiControlTimeResults[["36Cvs35C"]]$log2FoldChange,
#                               Fungi37Cvs36C = fungiControlTimeResults[["37Cvs36C"]]$log2FoldChange,
#                               Fungi38Cvs37C = fungiControlTimeResults[["38Cvs37C"]]$log2FoldChange,                             
#                               Fungi39Cvs38C = fungiControlTimeResults[["39Cvs38C"]]$log2FoldChange,
#                               Fungi40Cvs39C = fungiControlTimeResults[["40Cvs39C"]]$log2FoldChange,
#                               Fungi41Cvs40C = fungiControlTimeResults[["41Cvs40C"]]$log2FoldChange)

 
controlTimeList <- list(
  S20vs19 = spruceControlTimeResults[["20Cvs19C"]]$log2FoldChange,
  F20vs19 = fungiControlTimeResults[["20Cvs19C"]]$log2FoldChange,
  S21vs20 = spruceControlTimeResults[["21Cvs20C"]]$log2FoldChange,
  F21vs20 = fungiControlTimeResults[["21Cvs20C"]]$log2FoldChange,
  S22vs21 = spruceControlTimeResults[["22Cvs21C"]]$log2FoldChange,
  F22vs21 = fungiControlTimeResults[["22Cvs21C"]]$log2FoldChange,
  S23vs22 = spruceControlTimeResults[["23Cvs22C"]]$log2FoldChange,
  F23vs22 = fungiControlTimeResults[["23Cvs22C"]]$log2FoldChange,
  S24vs23 = spruceControlTimeResults[["24Cvs23C"]]$log2FoldChange,
  F24vs23 = fungiControlTimeResults[["24Cvs23C"]]$log2FoldChange,
  S25vs24 = spruceControlTimeResults[["25Cvs24C"]]$log2FoldChange,
  F25vs24 = fungiControlTimeResults[["25Cvs24C"]]$log2FoldChange,
  S26vs25 = spruceControlTimeResults[["26Cvs25C"]]$log2FoldChange,
  F26vs25 = fungiControlTimeResults[["26Cvs25C"]]$log2FoldChange,
  S28vs26 = spruceControlTimeResults[["28Cvs26C"]]$log2FoldChange,
  F28vs26 = fungiControlTimeResults[["28Cvs26C"]]$log2FoldChange,
  S31vs28 = spruceControlTimeResults[["31Cvs28C"]]$log2FoldChange,
  F31vs28 = fungiControlTimeResults[["31Cvs28C"]]$log2FoldChange,
  S32vs31 = spruceControlTimeResults[["32Cvs31C"]]$log2FoldChange,
  F32vs31 = fungiControlTimeResults[["32Cvs31C"]]$log2FoldChange,
  S34vs32 = spruceControlTimeResults[["34Cvs32C"]]$log2FoldChange,
  F34vs32 = fungiControlTimeResults[["34Cvs32C"]]$log2FoldChange,
  S35vs34 = spruceControlTimeResults[["35Cvs34C"]]$log2FoldChange,
  F35vs34 = fungiControlTimeResults[["35Cvs34C"]]$log2FoldChange,
  S36vs35 = spruceControlTimeResults[["36Cvs35C"]]$log2FoldChange,
  F36vs35 = fungiControlTimeResults[["36Cvs35C"]]$log2FoldChange,
  S37vs36 = spruceControlTimeResults[["37Cvs36C"]]$log2FoldChange,
  F37vs36 = fungiControlTimeResults[["37Cvs36C"]]$log2FoldChange,
  S38vs37 = spruceControlTimeResults[["38Cvs37C"]]$log2FoldChange,                             
  F38vs37 = fungiControlTimeResults[["38Cvs37C"]]$log2FoldChange,                             
  S39vs38 = spruceControlTimeResults[["39Cvs38C"]]$log2FoldChange,
  F39vs38 = fungiControlTimeResults[["39Cvs38C"]]$log2FoldChange,
  S40vs39 = spruceControlTimeResults[["40Cvs39C"]]$log2FoldChange,
  F40vs39 = fungiControlTimeResults[["40Cvs39C"]]$log2FoldChange,
  S41vs40 = spruceControlTimeResults[["41Cvs40C"]]$log2FoldChange,
  F41vs40 = fungiControlTimeResults[["41Cvs40C"]]$log2FoldChange
)
                      #FungiFvsC = resFungiTreatment$FertilisedvsControl$log2FoldChange)



```

```{r}
boxplot(controlTimeList,ylim=c(-2,2), xaxt = "n", col=c("#00FF00", "#FF0000"), ylab="log2FoldChange" )
text(x = 1:length(controlTimeList),
     ## Move labels to just below bottom of chart.
     y = par("usr")[3] - 0.45,
     ## Use names from the data list.
     labels = names(controlTimeList),
     ## Change the clipping region.
     xpd = NA,
     ## Rotate the labels by 35 degrees.
     srt = 90,
     ## Adjust the labels to almost 100% right-justified.
     adj = 0.365,
     ## Increase label size.
     cex = 1)

# ggplot(reshape2::melt(controlTimeList), aes(x=L1, y=value, col=L1  )) +
#   #geom_violin() +
#   #stat_summary(fun.y=mean, geom="point", shape=23, size=2) +
# 
#   geom_boxplot(width=0.1) +
#   stat_summary(fun.y=median, geom="point", size=2, color="black") +
#   ylim(-5, 5)
```

```{r}
ggplot(reshape2::melt(controlTimeList), aes(x=L1, y=value, col=L1  )) +
  #geom_violin() +
  #stat_summary(fun.y=mean, geom="point", shape=23, size=2) +

  geom_boxplot(width=0.1) +
  #stat_summary(fun.y=median, geom="point", size=2, color="black") +
  ylim(-2, 2)
```

## Fertilised time
```{r}
fertilisedTimeList <- list(
  S20vs19 = spruceFertilisedTimeResults[["20Fvs19F"]]$log2FoldChange,
  F20vs19 = fungiFertilisedTimeResults[["20Fvs19F"]]$log2FoldChange,
  S21vs20 = spruceFertilisedTimeResults[["21Fvs20F"]]$log2FoldChange,
  F21vs20 = fungiFertilisedTimeResults[["21Fvs20F"]]$log2FoldChange,
  S22vs21 = spruceFertilisedTimeResults[["22Fvs21F"]]$log2FoldChange,
  F22vs21 = fungiFertilisedTimeResults[["22Fvs21F"]]$log2FoldChange,
  S23vs22 = spruceFertilisedTimeResults[["23Fvs22F"]]$log2FoldChange,
  F23vs22 = fungiFertilisedTimeResults[["23Fvs22F"]]$log2FoldChange,
  S24vs23 = spruceFertilisedTimeResults[["24Fvs23F"]]$log2FoldChange,
  F24vs23 = fungiFertilisedTimeResults[["24Fvs23F"]]$log2FoldChange,
  S25vs24 = spruceFertilisedTimeResults[["25Fvs24F"]]$log2FoldChange,
  F25vs24 = fungiFertilisedTimeResults[["25Fvs24F"]]$log2FoldChange,
  S26vs25 = spruceFertilisedTimeResults[["26Fvs25F"]]$log2FoldChange,
  F26vs25 = fungiFertilisedTimeResults[["26Fvs25F"]]$log2FoldChange,
  S28vs26 = spruceFertilisedTimeResults[["28Fvs26F"]]$log2FoldChange,
  F28vs26 = fungiFertilisedTimeResults[["28Fvs26F"]]$log2FoldChange,
  S31vs28 = spruceFertilisedTimeResults[["31Fvs28F"]]$log2FoldChange,
  F31vs28 = fungiFertilisedTimeResults[["31Fvs28F"]]$log2FoldChange,
  S32vs31 = spruceFertilisedTimeResults[["32Fvs31F"]]$log2FoldChange,
  F32vs31 = fungiFertilisedTimeResults[["32Fvs31F"]]$log2FoldChange,
  S34vs32 = spruceFertilisedTimeResults[["34Fvs32F"]]$log2FoldChange,
  F34vs32 = fungiFertilisedTimeResults[["34Fvs32F"]]$log2FoldChange,
  S35vs34 = spruceFertilisedTimeResults[["35Fvs34F"]]$log2FoldChange,
  F35vs34 = fungiFertilisedTimeResults[["35Fvs34F"]]$log2FoldChange,
  S36vs35 = spruceFertilisedTimeResults[["36Fvs35F"]]$log2FoldChange,
  F36vs35 = fungiFertilisedTimeResults[["36Fvs35F"]]$log2FoldChange,
  S37vs36 = spruceFertilisedTimeResults[["37Fvs36F"]]$log2FoldChange,
  F37vs36 = fungiFertilisedTimeResults[["37Fvs36F"]]$log2FoldChange,
  S38vs37 = spruceFertilisedTimeResults[["38Fvs37F"]]$log2FoldChange,                             
  F38vs37 = fungiFertilisedTimeResults[["38Fvs37F"]]$log2FoldChange,                             
  S39vs38 = spruceFertilisedTimeResults[["39Fvs38F"]]$log2FoldChange,
  F39vs38 = fungiFertilisedTimeResults[["39Fvs38F"]]$log2FoldChange,
  S40vs39 = spruceFertilisedTimeResults[["40Fvs39F"]]$log2FoldChange,
  F40vs39 = fungiFertilisedTimeResults[["40Fvs39F"]]$log2FoldChange,
  S41vs40 = spruceFertilisedTimeResults[["41Fvs40F"]]$log2FoldChange,
  F41vs40 = fungiFertilisedTimeResults[["41Fvs40F"]]$log2FoldChange
)
```

```{r}
boxplot(fertilisedTimeList, ylim=c(-2,2), xaxt = "n", col=c("#00FF00", "#FF0000"), ylab="log2FoldChange" )
text(x = 1:length(fertilisedTimeList),
     ## Move labels to just below bottom of chart.
     y = par("usr")[3] - 0.45,
     ## Use names from the data list.
     labels = names(fertilisedTimeList),
     ## Change the clipping region.
     xpd = NA,
     ## Rotate the labels by 35 degrees.
     srt = 90,
     ## Adjust the labels to almost 100% right-justified.
     adj = 0.365,
     ## Increase label size.
     cex = 1)

```

# Session information
```{r}
sessionInfo()
```