---
title: "Spruce roots 2012 project - Differential expression analysis"
author: "Alonso Serrano"
date: "`r Sys.Date()`"
output:
 html_document:
   toc: true
   number_sections: true
---

# Introduction
This file contains the script and results for the differential expression analysis done to the
spruce roots Flakaliden 2012 data. 

*200324 Creation

## Prerequisites
All RNAseq Fastq files where previously processed by a nextflow pipeline for RNAseq analysis.
The files where QA with fastQC, then trimmed by Trimmomatic and sortmeRNA and finally aligned by
Salmon.
The output of the pipeline are the Salmon files, one per original fastq file.

A metadata file containing all the information related to each sample, file name, time, condition and replica number.

# Setup
Load libraries
```{r, message=FALSE}
library(DESeq2)
library(tximport)
library(BiocParallel)
library(plotly)
library(dplyr)
library(here)
register(MulticoreParam(4))
source(here("Rtoolbox/src/utilsDE.r"))
source(here("Rtoolbox/src/plot3Dvector.R"))
source(here("Rtoolbox/src/plotVectorPCA.R"))
source(here("Rtoolbox/src/plotUpAndDown.R"))
```

Set the project folder

```{r}
projectFolder <- "/mnt/picea/projects/spruce/vhurry/spruce-flakaliden-root-2012"
deFolder <- file.path(projectFolder, "DE")
```

```{r loadDEpreprocessed, echo=FALSE}
load(file.path(deFolder, "allDEresults.RData"))
```

Read metadata
```{r}
meta <- read.table(file.path(projectFolder, "data.csv"), header = TRUE, sep = ';', stringsAsFactors = T)
```


Add the interaction group to the metadata, in order to simplify
```{r}
meta$group <- paste0(meta$treatment, "_", meta$date)
meta$group <- factor(meta$group)
head(meta, n=3)

meta$date <- factor(meta$date, levels = c("Early_June", "Late_June", "August", "October") )
meta$treatment <- factor(meta$treatment, levels = c("Control", "5_year", "25_year") )
```

Get the Salmon files
```{r}
files <- dir(file.path(projectFolder, "salmon"),'quant.sf', recursive = TRUE, full.names = TRUE)
head(files)
```

Extract Sample ID from filenames and check if all files are present
```{r}
mvec <- basename(dirname((files)))
#mvec <- sub("(_S.*_L.*)", "", mvec)
if (! all(meta$SciLifeID %in% sub("(_S.*_L.*)", "", mvec)) ) {
  stop("Metadata doesn't match samples")
} else {
  print("All files present")
}

# order matching
all(sub("(_S.*_L.*)", "", mvec) == meta$SciLifeID)
```

Get the proper paths for the each Salmon quant.sf file
```{r}
files <- file.path(projectFolder, "salmon", mvec, "quant.sf")
```

Import transcript id to gene id table
```{r results="hide", message=FALSE}
tx2gene <- read.table(file.path(projectFolder,"tx2gene.tsv"), header = T, sep = '\t', stringsAsFactors = F)
txi <- tximport(files, type="salmon", tx2gene=tx2gene)
```

Import data as a DESeq object and place the sample names to the columns
```{r message=FALSE}
ddsMatrix <- DESeqDataSetFromTximport(txi, colData=meta, ~group)
colnames(ddsMatrix) <- meta$SciLifeID
```

check the DESeqDataSet
```{r}
ddsMatrix
```

Some genes have 0 expression accross all samples we can remove them
```{r}
ddsMatrix <- ddsMatrix[(rowSums(counts(ddsMatrix)) > 0),]
ddsMatrix
```

We save the gene names of all expressed genes for future uses
```{r}
# write.table(rownames(assay(ddsMatrix)), file="backgroundCombined.txt", sep='\t', row.names = F, col.names=F, quote = F)
```


# QA
## PCA
We use blind vst to plot a PCA and check for patterns and outliers

```{r}
vsd.QA <- vst(ddsMatrix, blind=TRUE)
geneNumber <- length(rownames(assay(vsd.QA)))
```
PCA using only treatment metadata
```{r}
plotPCA(vsd.QA, intgroup = c("treatment"), ntop=geneNumber)
```

PCA using only date metadata
```{r}
plotPCA(vsd.QA, intgroup = c("date"), ntop=geneNumber)
```

PCA using both treatment and date
```{r}
plotPCA(vsd.QA, intgroup = c("treatment", "date"), ntop=geneNumber)
```


The PCAs seems correct. 

## Focused PCA
We plot a PCA with letters instead of numbers for increased clarity
```{r}

vsd.QA.t <- t(assay(vsd.QA))
pca <- prcomp(vsd.QA.t)
percents <- round(summary(pca)$importance[2,]*100)
pchs <- LETTERS[as.integer(factor(meta$date))]
cols <- c("#f8766d", "#7cae00","#00bfc4")[as.integer(factor(meta$treatment))]
PCAlegend <- data.frame(pchs, meta$date)

par(oma = c(4, 1, 1, 4))
plot(pca$x[,1], pca$x[,2],
     xlab=paste("Comp. 1 (",percents[1],"%)",sep=""),
     ylab=paste("Comp. 2 (",percents[2],"%)",sep=""),
     col=cols,pch=pchs, cex = 1.0, main="")
par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
legend("bottomleft", legend=unique(PCAlegend$pchs:meta$date), cex = 0.8, xpd=TRUE)
legend("bottomright", legend=unique(meta$treatment), fill=unique(cols),cex = 0.8, xpd=TRUE)

```


## Scree plot
We can see how PC proportions are distributed
```{r}
screePlot(pca, 10)
```



The first 5 components explain 54 % of the variance
```{r}
summary(pca)$importance[2,][1:5] *100
sum(summary(pca)$importance[2,][1:5]) * 100
```



## PCA bi-plot comparition

As seen before, we can focus on the 5 components and plot a comparative bi-plot
```{r}
par(cex=1.0, cex.axis=0.8, cex.main=0.8)
pairs(pca$x[,1:5], col=cols, main="Principal components analysis bi-plot\nPCs 1-5", pch=pchs)
```



## 3D PCA
We represent the first 3 compoments as a 3D interactive object
```{r}
colors <- c("#f8766d", "#7cae00","#00bfc4")

treatment <- meta$treatment
# treatment[treatment == 'C'] <- 'Control'
# treatment[treatment == 'F'] <- 'Fertilised'

df3D <- data.frame(treatment)
df3D$time <- meta$date
df3D$colors <- colors[as.numeric(df3D$treatment)]
df3D$PC1 <- pca$x[,1]
df3D$PC2 <- pca$x[,2]
df3D$PC3 <- pca$x[,3] 

plot_ly(df3D, x = ~PC1, y = ~PC3, z = ~PC2, 
        color = ~treatment, colors = colors,
        text = ~paste("Date:", time, '<br>Treatment:', treatment) ) %>%   
  add_markers() %>%
  layout(scene = list(xaxis = list(title = paste("PC1 (",percents[1],"%)",sep="")),
                      yaxis = list(title = paste("PC3 (",percents[3],"%)",sep="")),
                      zaxis = list(title = paste("PC2 (",percents[2],"%)",sep=""))))

```



# Differential Expression
The dataset is ready for DESeq2. We check the design of the DE object 

```{r}
design(ddsMatrix)
```

We are interested on future comparitions like DE genes in each timepoints or between timepoints
that is why we created the interaction group

## Run DESeq
```{r runDE, eval=FALSE}
dds <- DESeq(ddsMatrix)
```

We check the result names
```{r}
resultsNames(dds)
```

In order to make all the comparitions automated, we prepare vectors with the elements that will be used
in the result extraction
```{r}
controlVector <- as.character(unique(meta$group[grep("Control", meta$group)]))
fertilised5Vector <- as.character(unique(meta$group[grep("^5_year", meta$group)]))
fertilised25Vector <- as.character(unique(meta$group[grep("25_year", meta$group)]))
```

## Result extraction
We want to make 3 different comparations, each one composed of several outputs from DESeq2
In order to do that, we will extract all the results for each comparation type, store them
as a list and then we can process them.

### Control date vs date
For this comparition we take consider only the control condition and we compare each pair of weeks
to determine which genes are up or down regaluated from one week to the previous week

```{r eval=FALSE}
control.res <- mapply(getRes, controlVector[-1],
                      controlVector[-length(controlVector)], 
                      MoreArgs = list(localDDS=dds, group="group"))
```

### Fertilised 5 date vs date
For this comparition we take consider only the fertilised condition and we compare each pair of weeks
to determine which genes are up or down regaluated from one week to the previous week

```{r eval=FALSE}
fertilised5.res <- mapply(getRes, fertilised5Vector[-1],
                          fertilised5Vector[-length(fertilised5Vector)], 
                          MoreArgs = list(localDDS=dds, group="group"))
```

### Fertilised 25 date vs date
For this comparition we take consider only the fertilised condition and we compare each pair of dates
to determine which genes are up or down regualted from one date to the previous date

```{r eval=FALSE}
fertilised25.res <- mapply(getRes, fertilised25Vector[-1], 
                         fertilised25Vector[-length(fertilised25Vector)],
                         MoreArgs = list(localDDS=dds, group="group"))
```

We add the corresponding names to the results
```{r eval=FALSE}
names(control.res) <- c("Control_Late_JunevsControl_Early_June", 
                        "Control_AugustvsControl_Late_June", 
                        "Control_OctobervsControl_August")
names(fertilised5.res) <- c("5_year_Late_Junevs5_year_Early_June", 
                        "5_year_Augustvs5_year_Late_June", 
                        "5_year_Octobervs5_year_August")
names(fertilised25.res) <- c("25_year_Late_Junevs25_year_Early_June", 
                        "25_year_Augustvs25_year_Late_June", 
                        "25_year_Octobervs25_year_August")
```

### Timepoint Fertilised 5 vs Control
For this comparition we take consider only a date each time and then compare DE genes due to 5 year fertilisation

```{r eval=FALSE}
combinedF5vsC.res <- mapply(getRes, fertilised5Vector, controlVector,
                            MoreArgs = list(localDDS=dds, group="group"))
```

### Timepoint Fertilised 25 vs Control
For this comparition we take consider only a date each time and then compare DE genes due 25 year  fertilisation

```{r eval=FALSE}
combinedF25vsC.res <- mapply(getRes, fertilised25Vector, controlVector,
                             MoreArgs = list(localDDS=dds, group="group"))
```

### Timepoint Fertilised 5 vs Control
For this comparition we take consider only a date each time and then compare DE genes due to 25 year fertilisation vs 5 year fertilisation

```{r eval=FALSE}
combinedF25vsF5.res <- mapply(getRes, fertilised25Vector, fertilised5Vector,
                              MoreArgs = list(localDDS=dds, group="group"))
```

We add the corresponding names to the results
```{r eval=FALSE}

names(combinedF5vsC.res) <- c("5_year_Early_JunevsControl_Early_June", 
                              "5_year_Late_JunevsControl_Late_June", 
                              "5_year_AugustvsControl_August", 
                              "5_year_OctobervsControl_October")
names(combinedF25vsC.res) <- c("25_year_Early_JunevsControl_Early_June", 
                               "25_year_Late_JunevsControl_Late_June", 
                               "25_year_AugustvsControl_August", 
                               "25_year_OctobervsControl_October")
names(combinedF25vsF5.res) <- c("25_year_Early_Junevs5_year_Early_June", 
                                "25_year_Late_Junevs5_year_Late_June", 
                                "25_year_Augustvs5_year_August", 
                                "25_year_Octobervs5_year_October")
```

## Filtering
The results contain all the genes, we need to filter them to get only those genes
that are stadistical significant and that have a strong condtition effect
For that we discard all the genes that has a padj higher than 0.01 and
a log2FoldChange lower than 0.5 (absolute value)
```{r filtering, results="hide", message=FALSE}
control.res.filter <- lapply(control.res, filterDE)
fertilised5.res.filter <- lapply(fertilised5.res, filterDE)
fertilised25.res.filter <- lapply(fertilised25.res, filterDE)
combinedF5vsC.res.filter <- lapply(combinedF5vsC.res, filterDE)
combinedF25vsC.res.filter <- lapply(combinedF25vsC.res, filterDE)
combinedF25vsF5.res.filter <- lapply(combinedF25vsF5.res, filterDE)
```

It is a good practice to do a check and see if the filtering was correct
```{r checkfilter, results="hide", message=FALSE}
lapply(control.res.filter, function(x) print(all(abs(x$log2FoldChange)>0.5 & x$padj<0.01)) )
lapply(fertilised5.res.filter, function(x) print(all(abs(x$log2FoldChange)>0.5 & x$padj<0.01)) )
lapply(fertilised25.res.filter, function(x) print(all(abs(x$log2FoldChange)>0.5 & x$padj<0.01)) )
lapply(combinedF5vsC.res.filter, function(x) print(all(abs(x$log2FoldChange)>0.5 & x$padj<0.01)) )
lapply(combinedF25vsC.res.filter, function(x) print(all(abs(x$log2FoldChange)>0.5 & x$padj<0.01)) )
lapply(combinedF25vsF5.res.filter, function(x) print(all(abs(x$log2FoldChange)>0.5 & x$padj<0.01)) )
```

## Plot DE genes
We can plot the results as positive and negative bars to reflect the amount of genes up or down
regulated in each comparition 

### Control date+1 vs date plot
```{r}
plotUpAndDown(control.res.filter)
```

### Fertilised 5 year date+1 vs date plot
```{r}
plotUpAndDown(fertilised5.res.filter)
```

### Fertilised 25 year date+1 vs date plot
```{r}
plotUpAndDown(fertilised25.res.filter)
```

### Fertilised 5 years vs Control per date plot
```{r}
plotUpAndDown(combinedF5vsC.res.filter)
```

### Fertilised 25 years vs Control per date plot
```{r}
plotUpAndDown(combinedF25vsC.res.filter)
```

### Fertilised 25 years vs Fertilised 5 years per date plot
```{r}
plotUpAndDown(combinedF25vsF5.res.filter)
```


# Results export
## Data for networks
For a co-expression network construction we need the vst data after doing a zeroMAD (Mean Decrease Accuracy) cleanup.
That will remove genes with zero or almost zero expression.
For the network construction with Seidr we need a headless files with the data and a file only with the gene names

### Combined VST data
```{r}
vsdCombined <- vst(dds, blind=FALSE)
combinedData <- t(assay(vsdCombined))
combinedData <- combinedData[,which(colMads(as.matrix(combinedData)) > 0)]
write.table(combinedData, file =file.path(deFolder, "spruceRoots2012Combined.tsv"), 
            sep='\t', row.names=FALSE, col.names=FALSE)
writeLines(paste(colnames(combinedData), collapse = '\t'),
           con=file.path(deFolder,"spruceRoots2012CombinedNames.txt"))
```


## DE results
For downstream analysis we export each set of results to a tsv file containing the padj and log2foldchange values for each comparition.
We export the results unfiltered
```{r exportDEresults, results="hide", message=FALSE}
de4network(control.res, file.path(deFolder,"deControl.tsv"))
de4network(fertilised5.res, file.path(deFolder,"deFertilised5.tsv"))
de4network(fertilised25.res, file.path(deFolder,"deFertilised25.tsv"))

de4network(combinedF5vsC.res, file.path(deFolder,"deCombinedF5vsC.tsv"))
de4network(combinedF25vsC.res, file.path(deFolder,"deCombinedF25vsC.tsv"))
de4network(combinedF25vsF5.res, file.path(deFolder,"deCombinedF25vsF5.tsv"))
```



# Session information
```{r}
sessionInfo()
```
```{r saveresults,eval=FALSE, echo=FALSE}

save(dds, control.res, fertilised5.res, fertilised25.res, 
     combinedF5vsC.res, combinedF25vsF5.res, combinedF25vsC.res,
     file=file.path(deFolder, "allDEresults.RData"))
```




# References
