---
title: "Network Analysis"
author: "Alonso Serrano"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    number_sections: yes
    toc: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction
This file contains the script and results of the network clustering of the Spruce 2011 root dataset. 

## Prerequisites
The variance stabilization transformed dataset, either as TSV or RData obtained from de differential expression analysis 
step.
Seidr aggregated network file that has a threshold applied


## Setup

```{r ini, message=FALSE, warning=FALSE}
library(RLinuxModules)
library(data.table)
module("load bioinfo-tools seidr-devel")
module("load bioinfo-tools InfoMap")
source("~/Rtoolbox/infomapTools.R")
source("~/Rtoolbox/plotEigenGene.R")
source("~/Rtoolbox/plotEnrichedTreemap.R")
source('~/Git/UPSCb/src/R/gopher.R')
seidrExe <- ("/mnt/picea/home/bastian/Git/seidr-devel/build/seidr ")
projectDir <- "/mnt/picea/projects/spruce/vhurry/14_SpruceRoots_Project/"
workDir <- paste0(projectDir,"spruce-networks/")
deDir <- paste0(projectDir,"DE/")
background <- read.table(paste0(deDir, "background.txt"), stringsAsFactors = F, header=F)[[1]]
load(paste0(deDir, "controlVsdData.RData"))
load(paste0(deDir, "fertilisedVsdData.RData"))
load(paste0(deDir, "combinedVsdData.RData"))
```
## Auxilary functions
```{r , message=FALSE, warning=FALSE}
plotResults <- function(clusterDir, theData, theClusters, theEnrichment,colors = c("#009444", "#BE9230")) {
  #crete the dir if it doesn't exist
  dir.create(clusterDir, showWarnings = FALSE)
  
  #save  results
  for (clus in 1:15) {
    
    #clusName <- names(theClusters)[clus]
    clusName <- paste0('Cluster',clus)
    #print(clusName)
    
    
    png(paste0(clusterDir, clusName, ".png" ), width = 1482, height=571 )
    print(plotEigengene(theData, theClusters[[clusName]], 
                        substr(rownames(theData),5,5), 
                        as.integer(substr(rownames(theData),2,3)), 
                        timeUnits = "Week",
                        colors=colors#,
                        #noLegend = F,
                        #legendTitle="Condition",
                        #title = paste(clusName,"Genes:", length(theClusters[[clusName]]) )
                        ))
    dev.off()
    
    png(paste0(clusterDir, clusName,"_enr_","go_bp",".png" ), width = 1482, height=571 )
    enr <- theEnrichment[[clusName]][['go']]

    if (!is.null(enr) && any(enr$namespace =='BP')) {
      plotEnrichedTreemap(theEnrichment[[clusName]], enrichment = 'go', 
                        namespace = 'BP',
                        clusterColor=clusterTreemapColors[clus],
                        clusterText=clusterTreemapText[clus])
    }
    dev.off()
      png(paste0(clusterDir,clusName,"_enr_","mapman",".png" ), width = 1482, height=571 )
    enr <- theEnrichment[[clusName]][['mapman']]

    if (!is.null(enr)) {
      plotEnrichedTreemap(theEnrichment[[clusName]], enrichment = 'mapman', 
                        clusterColor=clusterTreemapColors[clus],
                        clusterText=clusterTreemapText[clus])
    }
    dev.off()
  }
   
  
  for (clus in 1:15) {
  
    clusName <- paste0('Cluster',clus)
    cat(paste0("<H3>", clusName, "","</H3>"))
    cat("\n\n")
    print(plotEigengene(theData, theClusters[[clusName]],                      
                  substr(rownames(theData),5,5), 
                        as.integer(substr(rownames(theData),2,3)), 
                        timeUnits = "Week",
                        noLegend = F,
                        legendTitle="Condition",
                    colors=colors,
                        title = paste(clusName,"Genes:", length(theClusters[[clusName]]) )))
     cat("\n")
    enr <- theEnrichment[[clusName]][['go']]
    if (!is.null(enr) && any(enr$namespace =='BP')) {
      
      plotEnrichedTreemap(theEnrichment[[clusName]], enrichment = 'go', namespace = 'BP',
                          clusterColor=clusterTreemapColors[clus],
                          clusterText=clusterTreemapText[clus], 
                          title = paste(clusName," GO enrichment for BP"))
    }
    cat("\n")
    cat("<hr>")
        enr <- theEnrichment[[clusName]][['mapman']]

    if (!is.null(enr)) {
      plotEnrichedTreemap(theEnrichment[[clusName]], enrichment = 'mapman', 
                        clusterColor=clusterTreemapColors[clus],
                        clusterText=clusterTreemapText[clus], 
                          title = paste(clusName," Mapman enrichment for BP"))
    }

    cat("\n")
    cat("<hr>")
  }
  
}
```


# Network clustering
For each network, we will first some parameters, like folder location of the Seidr file, the markov-time parameter we found acceptable for clustering, the index edge file obtained from Seidr.

## Control network
### Clustering
First we define the parameters for the control network, the location of the seidr file and the markov time
```{r control}
netSubDir <- paste0(workDir, "control/aggregated/")
seidrFile <- paste0(netSubDir, "aggregated_0.28122.sf")
clusterFolder <- paste0(workDir, "control/cluster/")

edgeIndexFile <- paste0(netSubDir, "controlIndexEdgeList.txt")
treeFile <- paste0(clusterFolder,"controlIndexEdgeList.tree")
markovTime <- 0.54

dir.create(clusterFolder, showWarnings = FALSE)

system(paste("Infomap ", edgeIndexFile," -z --markov-time ", markovTime," ", clusterFolder))
infomapRes <- system(paste0(seidrExe, " resolve -s ", seidrFile, " ", treeFile), intern=TRUE)
infomapTable <-data.frame(do.call(rbind, strsplit(infomapRes, "\t")))
infomapTable <- prepareData(infomapTable)
infomapTable$Level1 <- infomapTable$P1
infomapTable$Level2 <- paste0(infomapTable$Level1,":",infomapTable$P2)
infomapTable$Level2 <- ifelse(infomapTable$Level2 %like% "NA", NA, infomapTable$Level2)  

clusterQA(infomapTable)
clusterQA(infomapTable, level='Level2')
```

```{r message=FALSE, warning=FALSE, include=FALSE}
minSize <- 200
selectedClusters <- getClusterByMinSize(infomapTable, min=minSize)
controlClusters <- getClusters(infomapTable, "Level1", numberOfClusters=selectedClusters)
save4Cytoscape(controlClusters, file=paste0(clusterFolder,"InfomapClusters.tsv"))
```

### Enrichment
```{r, eval=F}
controlEnr <- enrichClusters(controlClusters, background = background)
```
```{r loadEnr, include=FALSE}
#save(controlEnr, file=paste0(clusterFolder,"controlEnr.RData"))
load(file=paste0(clusterFolder, "controlEnr.RData"))
#print(controlEnr)
```

### Plot results
```{r plotControl, message=FALSE, warning=FALSE, results='asis'}
# clusterDir <- paste0(clusterFolder, "controlClustersPlots/")
# theData <- controlData
# theClusters <- controlClusters
# theEnrichment <- controlEnr

plotResults(paste0(clusterFolder, "controlClustersPlots/"), controlData, controlClusters, controlEnr)


```

## Fertilised network
### Clustering
```{r fertilised}
netSubDir <- paste0(workDir, "fertilised/aggregated/")
seidrFile <- paste0(netSubDir, "aggregated_0.26794.sf")
clusterFolder <- paste0(workDir, "fertilised/cluster/")

edgeIndexFile <- paste0(netSubDir, "fertilisedIndexEdgeList.txt")
treeFile <- paste0(clusterFolder,"fertilisedIndexEdgeList.tree")
markovTime <- 0.5

dir.create(clusterFolder, showWarnings = FALSE)

system(paste("Infomap ", edgeIndexFile," -z --markov-time ", markovTime," ", clusterFolder))
infomapRes <- system(paste0(seidrExe, " resolve -s ", seidrFile, " ", treeFile), intern=TRUE)
infomapTable <-data.frame(do.call(rbind, strsplit(infomapRes, "\t")))
infomapTable <- prepareData(infomapTable)
infomapTable$Level1 <- infomapTable$P1
infomapTable$Level2 <- paste0(infomapTable$Level1,":",infomapTable$P2)
infomapTable$Level2 <- ifelse(infomapTable$Level2 %like% "NA", NA, infomapTable$Level2)  

clusterQA(infomapTable)
clusterQA(infomapTable, level='Level2')
```

```{r}
minSize <- 200
selectedClusters <- getClusterByMinSize(infomapTable, min=minSize)
fertilisedClusters <- getClusters(infomapTable, "Level1", numberOfClusters=selectedClusters)
save4Cytoscape(fertilisedClusters, file=paste0(clusterFolder,"InfomapClusters.tsv"))
```

### Enrichment
```{r enrFertilised, eval=F}
fertilisedEnr <- enrichClusters(fertilisedClusters, background = background)
```
```{r loadfertEnr, include=FALSE}
#save(fertilisedEnr, file=paste0(clusterFolder,"fertilisedEnr.RData"))
load(file=paste0(clusterFolder, "fertilisedEnr.RData"))
#print(fertilisedEnr)
```


### Plot results
```{r plotFertilised, results='asis'}
plotResults(paste0(clusterFolder, "fertilisedClustersPlots/"), fertilisedData, fertilisedClusters, fertilisedEnr, colors = c("#BE9230","#009444"))


```


## Combined network
### Clustering
```{r combined}
netSubDir <- paste0(workDir, "combined/aggregated/")
seidrFile <- paste0(netSubDir, "aggregated_0.27043.sf")
clusterFolder <- paste0(workDir, "combined/cluster/")

edgeIndexFile <- paste0(netSubDir, "combinedIndexEdgeList.txt")
treeFile <- paste0(clusterFolder,"combinedIndexEdgeList.tree")
markovTime <- 0.53

dir.create(clusterFolder, showWarnings = FALSE)

system(paste("Infomap ", edgeIndexFile," -z --markov-time ", markovTime," ", clusterFolder))
infomapRes <- system(paste0(seidrExe, " resolve -s ", seidrFile, " ", treeFile), intern=TRUE)
infomapTable <-data.frame(do.call(rbind, strsplit(infomapRes, "\t")))
infomapTable <- prepareData(infomapTable)
infomapTable$Level1 <- infomapTable$P1
infomapTable$Level2 <- paste0(infomapTable$Level1,":",infomapTable$P2)
infomapTable$Level2 <- ifelse(infomapTable$Level2 %like% "NA", NA, infomapTable$Level2)  

clusterQA(infomapTable)
clusterQA(infomapTable, level='Level2')
```

```{r}
minSize <- 250
selectedClusters <- getClusterByMinSize(infomapTable, min=minSize)
combinedClusters <- getClusters(infomapTable, "Level1", numberOfClusters=selectedClusters)
save4Cytoscape(combinedClusters, file=paste0(clusterFolder,"InfomapClusters.tsv"))
```

### Enrichment
```{r, eval=F}
combinedEnr <- enrichClusters(combinedClusters, background = background)
```
```{r loadCombEnr, include=FALSE}
#save(combinedEnr, file=paste0(clusterFolder,"combinedEnr.RData"))
load(file=paste0(clusterFolder, "combinedEnr.RData"))
#print(combinedEnr)
```

### Plot results
```{r plotCombined, results='asis'}
plotResults(paste0(clusterFolder, "combinedClustersPlots/"), combinedData, combinedClusters, combinedEnr)


```
# Session information
Seidr version 0.11.0

Infomap version 0.19.26


```{r}
sessionInfo()
```
