---
title: "Fungi 2011 project - Differential expression analysis for flakaliden 2011 fungi data"
author: "Alonso Serrano"
date: "`r Sys.Date()`"
output:
 html_document:
   toc: true
   number_sections: true
---

# Introduction
This file contains the script and results for the differential expression analysis done to the
fungi 2011 data..

* 20191220 Updated with 2019-12-20 files from John  
* 

## Prerequisites
Transcript coutns file or megaData

A metadata file containing all the information related to each sample, file name, time, condition and replica number.


# Setup
Load libraries
```{r, message=FALSE}
library(DESeq2)
library(tximport)
library(BiocParallel)
library(plotly)
library(dplyr)
library(here)
register(MulticoreParam(4))
source(here("Rtoolbox/src/utilsDE.r"))
source(here("/Rtoolbox/src/plotInteractivePCA.R"))
source(here("/Rtoolbox/src/plotVectorPCA.R"))
source(here("Rtoolbox/src/plotEnrichedTreemap.R"))
source(here("Rtoolbox/src/plotUpAndDown.R"))
source(here("UPSCb-common/src/R/gopher.R"))

```
```{r}
plotResults <- function(theData) {
  
  for (deName in names(theData)) {
  
    #deName <- paste0('Cluster',clus)
    #clusIndex <- match(deName, names(theClusters))
    #clusIndex <- as.numeric(substr(deName,8,nchar(deName)))
    cat(paste0("<H3>", deName, "","</H3>"))
    cat("\n\n")
    
    enrUp <- theData[[deName]]$up[['ko_pathway']]
    enrDown <- theData[[deName]]$down[['ko_pathway']]
    
    cat("Up genes \n\n")
    if (!is.null(enrUp) ) {
      plotEnrichedTreemap(theData[[deName]]$up, enrichment = 'ko_pathway', namespace = 'none',
                          de='up', title = paste(deName," KO enrichment"))
    } else {
      cat("No enrichment data found")
    }
    cat("\n")
    cat("<hr>")
    
    cat("Down genes \n\n")

    
    if (!is.null(enrDown) ) {
      plotEnrichedTreemap(theData[[deName]]$down, enrichment = 'ko_pathway', namespace = 'none',
                          de='down', title = paste(deName," KO enrichment"))
    } else {
      cat("No enrichment data found")
    }
    cat("\n")
    cat("<hr>")

  }
  
}

getDEenrichment <- function(deData, background){
  results <- lapply(deData, function(x){

    upGenes <- rownames(x[x$log2FoldChange>0,])
    downGenes <- rownames(x[x$log2FoldChange<0,])
    upEnr <- list(ko_pathway=NULL)
    downEnr <- list(ko_pathway=NULL)
    
    if(length(upGenes)>1) 
      upEnr <- gopher(upGenes, task=list("ko_pathway"), background=background, url='ko', alpha=0.05)

    if(length(downGenes)>1) 
      downEnr <- gopher(downGenes, task=list("ko_pathway"), background=background, url='ko', alpha=0.05)
    
    res <- list(up=upEnr, down=downEnr)
    
    res
  })
  
  results
}
```


Set the project folder

```{r}
projectFolder <- "/mnt/picea/projects/spruce/vhurry/fungi-flakaliden-2011"
deFolder <- file.path(projectFolder,"DE")
dataFolder <- file.path(projectFolder,"data")
load(file.path(dataFolder, "megaData.RData") )


data <- megaData[, c(1:112)]
rownames(data) <- data$gene
data <- data[,-1]

```

Read metadata
```{r}
meta <- data.frame(week=substr(colnames(data), 2, 3),
                   treatment=substr(colnames(data), 4, 4),
                   group=substr(colnames(data), 2, 4))
```


Add the interaction group to the metadata, in order to simplify
```{r}
head(meta, n=3)
```


Import data as a DESeq object and place the sample names to the columns
```{r deImport, message=FALSE}
ddsMatrix <- DESeqDataSetFromMatrix(data, colData=meta, design = ~group)
ddsMatrix
```


Remove 0 expression genes accross all samples we can remove it
```{r}
ddsMatrix <- ddsMatrix[(rowSums(counts(ddsMatrix)) > 0),]
ddsMatrix
```


# QA
## PCA
We use blind vst to plot a PCA and check for patterns and outliers
```{r vsdBlind}
vsd.QA <- vst(ddsMatrix, blind=TRUE, fitType = 'local')
```

Interactive PCA 
```{r message=FALSE, warning=FALSE}
plotInteractivePCA(t(assay(vsd.QA)),
          meta=data.frame(week=vsd.QA$week, treatment=vsd.QA$treatment), 
          group=c("treatment"),
          colors=c("#009444","#BE9230") )
```

# Differential Expression
The dataset is ready for DESeq2. We check the design of the DE object 

```{r}
design(ddsMatrix)
```

We are interested on future comparitions like DE ko in each timepoints or between timepoints
that is why we created the interaction group

## Run DESeq
```{r runDE, eval=T}
dds <- DESeq(ddsMatrix)
```

## Data for networks
For a co-expression network construction we need the vst data after doing a zeroMAD (Mean Decrease Accuracy) cleanup.
That will remove genes with zero or almost zero expression.
For the network construction with Seidr we need a headless files with the data and a file only with the gene names

### Combined
```{r,eval=FALSE}
vsdCombined <- vst(dds, blind=FALSE)
combinedData <- t(assay(vsdCombined))
combinedData <- combinedData[,which(colMads(as.matrix(combinedData)) > 0)]
save(combinedData, file=file.path(deFolder,"transcriptsCombinedVsdData.RData"))
write.table(combinedData, file =file.path(deFolder,"transcriptsFungi2011Combined.tsv"), sep='\t', row.names=FALSE, col.names=FALSE)
writeLines(colnames(combinedData), con=file.path(deFolder,"transcriptsFungi2011CombinedNames.txt"))
```


We check the result names
```{r}
resultsNames(dds)
```

In order to make all the comparitions automated, we prepare vectors with the elements that will be used
in the result extraction
```{r}
controlVector <- as.character(unique(meta$group[grep("C", meta$group)]))
fertilisedVector <- as.character(unique(meta$group[grep("F", meta$group)]))
```

## Result extraction
We want to make 3 different comparations, each one composed of several outputs from DESeq2
In order to do that, we will extract all the results for each comparation type, store them
as a list and then we can process them.

### Control week vs week
For this comparition we take consider only the control condition and we compare each pair of weeks
to determine which genes are up or down regaluated from one week to the previous week

```{r eval=FALSE}
control.res <- mapply(getRes, controlVector[-1],
                      controlVector[-length(controlVector)],
                      MoreArgs = list(localDDS=dds, group="group"))
```

### Fertilised week vs week
For this comparition we take consider only the fertilised condition and we compare each pair of weeks
to determine which genes are up or down regaluated from one week to the previous week

```{r eval=FALSE}
fertilised.res <- mapply(getRes, fertilisedVector[-1], 
fertilisedVector[-length(fertilisedVector)], 
MoreArgs = list(localDDS=dds, group="group"))
```

### Timepoint Fertilised vs Control
For this comparition we take consider only a week each time and then compare DE genes due to fertilisation

```{r eval=FALSE}
combined.res <- mapply(getRes, fertilisedVector, controlVector, 
MoreArgs = list(localDDS=dds, group="group"))
```

We add the corresponding names to the results
```{r addNames}
names(control.res) <- c("20Cvs19C", "21Cvs20C", "22Cvs21C", "23Cvs22C",
                     "24Cvs23C", "25Cvs24C", "26Cvs25C", "28Cvs26C",
                     "31Cvs28C", "32Cvs31C", "34Cvs32C", "35Cvs34C",
                     "36Cvs35C", "37Cvs36C", "38Cvs37C", "39Cvs38C",
                     "40Cvs39C", "41Cvs40C")

names(fertilised.res) <- c("20Fvs19F", "21Fvs20F", "22Fvs21F", "23Fvs22F",
                        "24Fvs23F", "25Fvs24F", "26Fvs25F", "28Fvs26F",
                        "31Fvs28F", "32Fvs31F", "34Fvs32F", "35Fvs34F",
                        "36Fvs35F", "37Fvs36F", "38Fvs37F", "39Fvs38F",
                        "40Fvs39F", "41Fvs40F")

names(combined.res) <- c("19Fvs19C", "20Fvs20C", "21Fvs21C", "22Fvs22C",
                      "23Fvs23C", "24Fvs24C", "25Fvs25C", "26Fvs26C",
                      "28Fvs28C", "31Fvs31C", "32Fvs32C", "34Fvs34C",
                      "35Fvs35C", "36Fvs36C", "37Fvs37C", "38Fvs38C",
                      "39Fvs39C", "40Fvs40C", "41Fvs41C")
```

```{r include=FALSE, eval=FALSE}
save(dds, control.res, fertilised.res, combined.res, file=file.path(deFolder,"allDEresults.RData"))
```

```{r include=FALSE}
load(file.path(deFolder,"allDEresults.RData"))
```

## Filtering
The results contain all the genes, we need to filter them to get only those genes
that are stadistical significant and that have a strong condtition effect
For that we discard all the genes that has a padj higher than 0.01 and
a log2FoldChange lower than 0.5 (absolute value)
```{r filtering, results="hide", message=FALSE}
control.res.filter <- lapply(control.res, filterDE)
fertilised.res.filter <- lapply(fertilised.res, filterDE)
combined.res.filter <- lapply(combined.res, filterDE)
```

It is a good practice to do a check and see if the filtering was correct
```{r checkfilter, results="hide", message=FALSE}
lapply(control.res.filter, function(x) print(all(abs(x$log2FoldChange)>0.5 & x$padj<0.01)) )
lapply(fertilised.res.filter, function(x) print(all(abs(x$log2FoldChange)>0.5 & x$padj<0.01)) )
lapply(combined.res.filter, function(x) print(all(abs(x$log2FoldChange)>0.5 & x$padj<0.01)) )
```

## Plot DE genes
We can plot the results as positive and negative bars to reflect the amount of genes up or down
regulated in each comparition 
For that we create a couple more vector to present the results

In order to get the results by week, and then treatment. 19C, 19F, 20C, ...
```{r}
mixVector <- c(rbind(controlVector, fertilisedVector))
```
And to do the same by treatment. 19C, 20C, ..., 41C, 19F, 20F, ...

```{r}
concatVector <- c(cbind(controlVector, fertilisedVector))
```

### Control week+1 vs week plot
```{r}
plotUpAndDown(control.res.filter)
```

### Fertilised week+1 vs week plot
```{r}
plotUpAndDown(fertilised.res.filter)
```

### Fertilised vs Control per week plot
```{r}
plotUpAndDown(combined.res.filter)
```

## DE Fertilisation vs Control
```{r}
design(ddsMatrix) <- ~Treatment

ddsFungiTreatment  <- DESeq(ddsMatrix)

resFungiTreatment <- list( FertilisedvsControl = results(ddsFungiTreatment ,
        name= "Treatment_F_vs_C",
        filter = rowMedians(counts(ddsFungiTreatment )),
        parallel = FALSE) 
        )

save(ddsFungiTreatment, resFungiTreatment, file=file.path(projectFolder, "DE", "ddsTreatment.RData"))
de4network(resFungiTreatment, file.path(deFolder,"deFvsC.tsv"))
resFungiTreatment.filter <- lapply(resFungiTreatment, filterDE)

plotUpAndDown(resFungiTreatment.filter)

```

## Top genes
To check the top genes in each comparition we create a dot plot for the top 6 genes with two lines (read and blue)
that highligths the comparition pair that we want to check and store it as a PDF

```{r plotTops, results="hide", message=FALSE,eval=FALSE}
plotTop(control.res.filter, dds, plotGroup=c('Treatment','Week'), infoVector=concatVector, name = file.path(deFolder,"controlFilter.pdf"))
plotTop(fertilised.res.filter, dds, plotGroup=c('Treatment','Week'), infoVector=concatVector, name = file.path(deFolder,"fertilisedFilter.pdf"))
plotTop(combined.res.filter, dds, plotGroup=c('Week','Treatment'), infoVector=mixVector, name = file.path(deFolder,"combinedFilter.pdf"))
```

# Results export
## Background for enrichment
We need to extract the names of all the genes that at least have 1 count across all samples to establish a background 
needed for future enrichment steps
```{r}
write.table(colnames(t(assay(dds))), file=file.path(deFolder,"background.txt"), row.names = F,col.names = F, quote = F )
```


## Data for networks
For a co-expression network construction we need the vst data after doing a zeroMAD (Mean Decrease Accuracy) cleanup.
That will remove genes with zero or almost zero expression.
For the network construction with Seidr we need a headless files with the data and a file only with the gene names

### Combined
```{r,eval=FALSE}
vsdCombined <- vst(dds, blind=FALSE)
combinedData <- t(assay(vsdCombined))
combinedData <- combinedData[,which(colMads(as.matrix(combinedData)) > 0)]
save(combinedData, file=file.path(deFolder,"combinedVsdData.RData"))
write.table(combinedData, file =file.path(deFolder,"fungi2011Combined.tsv"), sep='\t', row.names=FALSE, col.names=FALSE)
writeLines(colnames(combinedData), con=file.path(deFolder,"fungi2011CombinedNames.txt"))
```

### Control & fertilised
In order to make this individual networks, we need to specify a new design to obtain the vst data using model awareness. In this case, the model should be treatment.
```{r ,eval=FALSE}
ddsSplit <- dds
design(ddsSplit) <- ~Treatment
vsdSplit <- vst(ddsSplit, blind=F)
vsdSplitData <- t(assay(vsdSplit))
controlData <- vsdSplitData[grep("C",rownames(vsdSplitData)),]
controlData <- controlData[,which(colMads(as.matrix(controlData)) > 0)]

fertilisedData <- vsdSplitData[grep("F",rownames(vsdSplitData)),]
fertilisedData <- fertilisedData[,which(colMads(as.matrix(fertilisedData)) > 0)]

save(controlData, file=file.path(deFolder,"controlVsdData.RData"))
write.table(controlData, file = file.path(deFolder,"fungi2011Control.tsv"), sep='\t', row.names=FALSE, col.names=FALSE)
writeLines(colnames(controlData), con=file.path(deFolder,"fungi2011ControlNames.txt"))

save(fertilisedData, file=file.path(deFolder,"fertilisedVsdData.RData"))
write.table(fertilisedData, file = file.path(deFolder,"fungi2011Fertilised.tsv"), sep='\t', row.names=FALSE, col.names=FALSE)
writeLines(colnames(fertilisedData), con=file.path(deFolder,"fungi2011FertilisedNames.txt"))

```

## DE results
For downstream analysis we export each set of results to a tsv file containing the padj and log2foldchange values for each comparition.
We export the results unfiltered
```{r exportDEresults, eval=FALSE, message=FALSE, include=FALSE, results="hide"}
de4network(control.res, file.path(deFolder,"deControl.tsv"))
de4network(fertilised.res, file.path(deFolder,"deFertilised.tsv"))
de4network(combined.res, file.path(deFolder,"deCombined.tsv"))
```

## DE enrichment

We set the background
```{r}
background <- read.table(file.path(deFolder, "background.txt"), sep='\t', header=F, stringsAsFactors = F)[,1] 
```
Now we use gopher to obtain the enrichment results for up and down expressed genes
```{r}
control.res.enr <- getDEenrichment(control.res.filter, background)
fertilised.res.enr <- getDEenrichment(fertilised.res.filter, background)
combined.res.enr <- getDEenrichment(combined.res.filter, background)
```

### Control
```{r}
plotResults(control.res.enr)
```
### Fertilised
```{r}
plotResults(fertilised.res.enr)
```

### Combined
```{r}
plotResults(combined.res.enr)
```



# Session information
```{r}
sessionInfo()
```




# References
