---
title: "funGuild"
author: "Alonso Serrano"
date: "`r Sys.Date()`"
output:
 html_document:
   toc: true
   number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script contains the necessary steps to obtain the guild for the fungal taxo using the FUNGuild tool

Nguyen NH, Song Z, Bates ST, Branco S, Tedersoo L, Menke J, Schilling JS, Kennedy PG. 2016. FUNGuild: An open annotation tool for parsing fungal community datasets by ecological guild. Fungal Ecology 20:241–248.

*2020-03-23: added the Archaeorhizomcyete data


# Load fungi data
```{r load}
projectFolder <- "/mnt/picea/projects/spruce/vhurry/fungi-flakaliden-2011"
deFolder <- file.path(projectFolder, "DE")
dataFolder <- file.path(projectFolder, "data")

taxonomyFile <- file.path(dataFolder, "gene_taxonomy_update.tsv")
taxo <- read.table(taxonomyFile, header = TRUE, sep = '\t', comment.char = "", stringsAsFactors = FALSE)
```

# Export data for funGuild
```{r export}

taxo_df <- data.frame(gene = taxo$gene, taxonomy =apply(taxo[,2:8], 1, paste, collapse = ";" ) )
write.table(taxo_df, file=file.path(dataFolder, "taxo4funguild.tsv"), quote=F, row.names=F, sep='\t')
```

# Execute funGuild

To execute funGuild, first we need to clone the repository located in 
https://github.com/UMNFuN/FUNGuild

The using python 3.6.8 and the file generated before we execute the following command

```
python Guilds_v1.1.py -otu taxo4funguild.tsv

Execution looks as follows
FunGuild v1.0 Beta<br>
Connecting with FUNGuild database ...<br>

Reading in the OTU table: 'taxo4funguild.tsv'

Searching the FUNGuild database...
10%
20%
30%
40%
50%
60%
70%
80%
90%
100%

Found 114299 matching taxonomy records in the database.
Dereplicating and sorting the result...

...
FunGuild tried to assign function to 323012 OTUs in 'taxo4funguild.tsv'.
FUNGuild made assignments on 176579 OTUs.
Result saved to 'taxo4funguild.guilds.txt'

Total calculating time: 23371.12 seconds.
```

```{r sessionInfo}
sessionInfo()
```


