---
title: "Fungi network analysis - Gloniaceae - Fertilised"
author: "Alonso Serrano"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    number_sections: yes
    toc: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction
This file contains the script and results of the network clustering of the fungi 2011 dataset, for the Gloniaceae family . 

## Prerequisites
The variance stabilization transformed dataset, either as TSV or RData obtained from de differential expression analysis step.
Seidr aggregated network file that has a threshold applied


## Setup

```{r ini, message=FALSE, warning=FALSE}
library(RLinuxModules)
library(data.table)
module("load bioinfo-tools seidr-devel")
module("load bioinfo-tools InfoMap")
source("~/Git/Rtoolbox/src/infomapTools.R")
source("~/Git/Rtoolbox/src/plotEigenGene.R")
source("~/Git/Rtoolbox/src/plotEnrichedTreemap.R")
source('~/Git/UPSCb-common/src/R/gopher.R')
seidrExe <- ("/mnt/picea/home/bastian/Git/seidr-devel/build/seidr ")
projectDir <- "/mnt/picea/projects/spruce/vhurry/fungi-flakaliden-2011/Gloniaceae"
workDir <- file.path(projectDir,"networks")
deDir <- file.path(projectDir,"DE")
networkDir <- file.path(projectDir,"networks")
background <- read.table(file.path(deDir, "background.txt"), stringsAsFactors = F, header=F)[[1]]

load(file.path(deDir, "fertilisedVsdData.RData"))

```
## Auxilary functions
```{r , message=FALSE, warning=FALSE}
plotResults <- function(theData, theClusters, theEnrichment,colors = c("#009444", "#BE9230")) {
  for (clusName in names(theClusters)) {

    #clusIndex <- as.numeric(substr(clusName,8,nchar(clusName)))
    clusIndex <-which(names(theClusters) %in% clusName)
    cat(paste0("<H3>", clusName, "","</H3>"))
    cat("\n\n")
    print(plotEigengene(theData, theClusters[[clusName]],                      
                  substr(rownames(theData),4,4), 
                        as.integer(substr(rownames(theData),2,3)), 
                        timeUnits = "Week",
                        noLegend = F,
                        legendTitle="Condition",
                    colors=colors,
                        title = paste(clusName,"Genes:", length(theClusters[[clusName]]) )))

    cat("\n")

    enr <- theEnrichment[[clusName]][['kog']]
    if (!is.null(enr) ) {
          cat(paste0("<H4>KOG</H4>"))
      plotEnrichedTreemap(theEnrichment[[clusName]], enrichment = 'kog', namespace = 'BP',
                          clusterColor=clusterTreemapColors[clusIndex],
                          clusterText=clusterTreemapText[clusIndex], 
                          title = paste(clusName,"KOG enrichment"))
          cat("\n")
    cat("<hr>")
    }

        cat("\n")

    enr <- theEnrichment[[clusName]][['cog']]
    if (!is.null(enr) ) {
          cat(paste0("<H4>COG</H4>"))
      plotEnrichedTreemap(theEnrichment[[clusName]], enrichment = 'cog', namespace = 'BP',
                          clusterColor=clusterTreemapColors[clusIndex],
                          clusterText=clusterTreemapText[clusIndex], 
                          title = paste(clusName," COG enrichment"))
          cat("\n")
    cat("<hr>")
        cat("\n")
    }

      enr <- theEnrichment[[clusName]][['ko']]
    if (!is.null(enr) ) {
      cat(paste0("<H4>KO</H4>"))

      plotEnrichedTreemap(theEnrichment[[clusName]], enrichment = 'ko', namespace = 'BP',
                          clusterColor=clusterTreemapColors[clusIndex],
                          clusterText=clusterTreemapText[clusIndex], 
                          title = paste(clusName," KO enrichment"))
      cat("\n")
      cat("<hr>")
      cat("\n")
    }


    enr <- theEnrichment[[clusName]][['ko_pathway']]
    if (!is.null(enr) ) {
      cat(paste0("<H4>KO pathway</H4>"))
      plotEnrichedTreemap(theEnrichment[[clusName]], enrichment = 'ko_pathway', namespace = 'BP',
                          clusterColor=clusterTreemapColors[clusIndex],
                          clusterText=clusterTreemapText[clusIndex], 
                          title = paste(clusName," KO pathway enrichment"))
      cat("\n")
      cat("<hr>")
    }
      cat("\n")
      cat("<hr>")

  }
  
}
```


# Network clustering
For each network, we will first some parameters, like folder location of the Seidr file, the markov-time parameter we found acceptable for clustering, the index edge file obtained from Seidr.

## Network
### Clustering
First we define the parameters for the control network, the location of the seidr file and the markov time
```{r control}
network = 'fertilised'
seidrFile = "aggregated_0.289.sf"
theData <- fertilisedData
markovTime <- 1.1
minSize <- 40
theColors <- c("#BE9230")

netSubDir <- file.path(networkDir, network, "results/aggregated")
seidrFile <- file.path(netSubDir, seidrFile)
clusterFolder <- file.path(networkDir, network, "cluster")

dir.create(clusterFolder, showWarnings = FALSE)
```
# Infomap clustering
First we use Seidr reheader function to drop not connected nodes and improve efficiency
```{r reheader, results=F, message='HIDE'}
system(paste("seidr reheader" , seidrFile), intern=TRUE)

```
We execute seidr view to extract the information of source node, target node and the score of irp
```{r include=FALSE}
edgeIndexFile <- file.path(clusterFolder, "indexEdgeList.txt")
edgeFile <- file.path(clusterFolder, "edgeList.txt")
treeFile <- file.path(clusterFolder,"indexEdgeList.tree")



headResult <- system(paste("seidr view", seidrFile, "-c -d $'\t' ", "| head -n 1"), intern=TRUE)
headResult <- unlist(strsplit(headResult, "\t"))

algoIndex <- grep("irp_score", headResult)
system(paste0("seidr view ", seidrFile, "  -d $'\t' | cut -f 1,2,",algoIndex, " > ",edgeFile), intern=TRUE)
system(paste0("seidr view ", seidrFile, " -N -d $'\t' | cut -f 1,2,",algoIndex," >",edgeIndexFile), intern=TRUE)

#clean edgeFiles from weird $ character at the end of line
#
system(paste("sed -i 's/\\$//g' ", edgeIndexFile),intern=TRUE) 
system(paste("sed -i 's/\\$//g' ", edgeFile),intern=TRUE) 
```

```{r clustering}
system(paste("Infomap ", edgeIndexFile," -z --markov-time ", markovTime," ", clusterFolder))
infomapRes <- system(paste0(seidrExe, " resolve -s ", seidrFile, " ", treeFile), intern=TRUE)
infomapTable <-data.frame(do.call(rbind, strsplit(infomapRes, "\t")))
infomapTable <- prepareData(infomapTable)
infomapTable$Level1 <- infomapTable$P1
infomapTable$Level2 <- paste0(infomapTable$Level1,":",infomapTable$P2)
infomapTable$Level2 <- ifelse(infomapTable$Level2 %like% "NA", NA, infomapTable$Level2)  

clusterQA(infomapTable)
clusterQA(infomapTable, level='Level2')
```


```{r }

selectedClusters <- getClusterByMinSize(infomapTable, level='Level1', min=minSize)
theClusters <- getClusters(infomapTable, "Level1", numberOfClusters=selectedClusters)

```

```{r message=FALSE, warning=FALSE, include=FALSE}
save4Cytoscape(theClusters, file=file.path(clusterFolder,"InfomapClusters.tsv"))
```

## Enrichment
```{r enrich, message=FALSE, warning=FALSE, include=TRUE, results=F}
background <- as.character(infomapTable$gene)

theEnrichment <- lapply(theClusters, function(x) {
  print(length(x))
  if(length(x) > 1)
    gopher(x, task = list("ko_pathway","ko","cog","kog"), 
           background = background, url="fungi2011", alpha = 0.05)
  else
    NULL
})

```

```{r}
save(theEnrichment, file=file.path(clusterFolder, "clusterEnr.RData"))
```

```{r, include=FALSE, eval=FALSE}
load(file.path(clusterFolder, "clusterEnr.RData"))
```

## Plot results
```{r plotControl, message=FALSE, warning=FALSE, results='asis'}
plotResults(theData, theClusters, theEnrichment,colors=theColors)

```


# Session information
Seidr version 0.12.0

Infomap version 0.19.26


```{r}
sessionInfo()
```
