---
title: "Fungi network analysis - Species - Clusterinf"
author: "Alonso Serrano"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document: 
    number_sections: yes
    toc: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction


## Prerequisites



## Setup

```{r ini, message=FALSE, warning=FALSE}
library(RLinuxModules)
library(data.table)
library(here)
library(tidyverse)
library(stringr)
library(reshape2)
module("load bioinfo-tools seidr-devel")
module("load bioinfo-tools InfoMap")
library(dplyr)
source(here("Rtoolbox/src/utilsDE.r"))
source(here("/Rtoolbox/src/plot3Dvector.R"))
source(here("/Rtoolbox/src/plotVectorPCA.R"))
source(here("Rtoolbox/src/plotEnrichedTreemap.R"))
source(here("Rtoolbox/src/plotUpAndDown.R"))
source(here("UPSCb-common/src/R/gopher.R"))
source(here("Rtoolbox/src/getEigengenes.R"))

source("~/Git/Rtoolbox/src/infomapTools.R")
source("~/Git/Rtoolbox/src/plotEigenGene.R")
source("~/Git/Rtoolbox/src/plotEnrichedTreemap.R")
source('~/Git/UPSCb-common/src/R/gopher.R')

#seidrExe <- ("/mnt/picea/home/bastian/Git/seidr-devel/build/seidr ")
projectDir <- "/mnt/picea/projects/spruce/vhurry/fungi-flakaliden-2011/fungi-species"
species <- c("cenococcum_geophilum","cortinarius_glaucopus","hyaloscypha_variabilis","hyaloscypha_bicolor","piloderma_croceum")
#species <- c("cortinarius_glaucopus")
conditions <- c("control", "combined")
# workDir <- file.path("networks")
# deDir <- file.path(projectDir,"DE")
# networkDir <- file.path(projectDir,"networks")
# background <- read.table(file.path(deDir, "background.txt"), stringsAsFactors = F, header=F)[[1]]
# load(file.path(deDir, "controlVsdData.RData"))
#load(file.path(deDir, "fertilisedVsdData.RData"))
#

#dataFolder<- file.path(projectDir, "data")
#resultsFolder<- file.path(projectFolder, "reports")

load(file.path("/mnt/picea/projects/spruce/vhurry/fungi-flakaliden-2011/data", "megaData.RData"))
```
## Auxilary functions
```{r , message=FALSE, warning=FALSE}
plotResults <- function(theData, theClusters, theEnrichment,colors = c("#009444", "#BE9230")) {
  for (clusName in names(theClusters)) {

    clusIndex <-which(names(theClusters) %in% clusName)
    cat(paste0("<H3>", clusName, "","</H3>"))
    cat("\n\n")
    print(plotEigengene(theData, theClusters[[clusName]],                      
                  substr(rownames(theData),4,4), 
                        as.integer(substr(rownames(theData),2,3)), 
                        timeUnits = "Week",
                        noLegend = F,
                        legendTitle="Condition",
                    colors=colors,
                        title = paste(clusName,"Genes:", length(theClusters[[clusName]]) )))

    cat("\n")

    enr <- theEnrichment[[clusName]][['kog']]
    if (!is.null(enr) ) {
          cat(paste0("<H4>KOG</H4>"))
      plotEnrichedTreemap(theEnrichment[[clusName]], enrichment = 'kog', namespace = 'BP',
                          clusterColor=clusterTreemapColors[clusIndex],
                          clusterText=clusterTreemapText[clusIndex], 
                          title = paste(clusName,"KOG enrichment"))
          cat("\n")
    cat("<hr>")
    }

        cat("\n")

    enr <- theEnrichment[[clusName]][['cog']]
    if (!is.null(enr) ) {
          cat(paste0("<H4>COG</H4>"))
      plotEnrichedTreemap(theEnrichment[[clusName]], enrichment = 'cog', namespace = 'BP',
                          clusterColor=clusterTreemapColors[clusIndex],
                          clusterText=clusterTreemapText[clusIndex], 
                          title = paste(clusName," COG enrichment"))
          cat("\n")
    cat("<hr>")
        cat("\n")
    }

      enr <- theEnrichment[[clusName]][['ko']]
    if (!is.null(enr) ) {
      cat(paste0("<H4>KO</H4>"))

      plotEnrichedTreemap(theEnrichment[[clusName]], enrichment = 'ko', namespace = 'BP',
                          clusterColor=clusterTreemapColors[clusIndex],
                          clusterText=clusterTreemapText[clusIndex], 
                          title = paste(clusName," KO enrichment"))
      cat("\n")
      cat("<hr>")
      cat("\n")
    }


    enr <- theEnrichment[[clusName]][['ko_pathway']]
    if (!is.null(enr) ) {
      cat(paste0("<H4>KO pathway</H4>"))
      plotEnrichedTreemap(theEnrichment[[clusName]], enrichment = 'ko_pathway', namespace = 'BP',
                          clusterColor=clusterTreemapColors[clusIndex],
                          clusterText=clusterTreemapText[clusIndex], 
                          title = paste(clusName," KO pathway enrichment"))
      cat("\n")
      cat("<hr>")
    }
      cat("\n")
      cat("<hr>")

  }
  
}
```

# Thresholding

```{r seidrDescribe}
seidrDescribe <- function(seidrFile){
  require(stringr)
  require(tidyverse)
  require(dplyr)
  
  #system("module load bioínfo-tools seidr-devel")
  com <- paste0("seidr describe -b 100 -q $(seq -s ',' 0.01 0.01 0.99) ", seidrFile)
  #print(com)
  result <- system(com, intern=TRUE)
  
  #result <- as.data.frame(result, stringsAsFactors = F)
  result <- str_split_fixed(result, "\t", 5) 
  result <- as.data.frame(result, stringsAsFactors = F)
  result <- result %>% filter(V2=='irp' & V3=='HIST') %>% select(V4,V5)
  names(result) <- c("range", "counts")
  
  # clean counts column
  result$counts <- gsub("#","",result$counts)
  result$counts <- gsub(" ","",result$counts)
  result$counts <- as.integer(result$counts)
  result <- result %>% map_df(rev)
  result <- result %>% mutate(cumCounts = cumsum(counts))
  
  total <- tail(result, 1)$cumCounts
  onePercent <- floor(total/100)
  
  idx <- which(result$cumCounts> onePercent)[1]
  selLine <- result[idx,]
  
  threshold <- as.numeric(str_extract(selLine$range, "([0-1]+\\.[0-9]*)"))
  
  return(threshold)
  
} 
```

```{r}
seidrThreshold <- function(seidrFile, min) {
  require(stringr)
  require(tidyverse)
  require(dplyr)
  folder <- dirname(seidrFile)
  outFile <- file.path(folder, "threshold.txt")
  capture <- system(paste("seidr threshold -n 1000 -m", min, "-M 1.0", seidrFile, "-f -o", outFile, "2>&1"), intern=TRUE)
  suggestedThreshold <- as.numeric(str_extract(tail(capture,1), "([0-1]+\\.[0-9]*)"))
  

  
  thresholdTable <- read.table(outFile, sep='\t', stringsAsFactors = F)
  names(thresholdTable) <- c("Threshold", "Nodes", "Edges", "SFT", "ACC")
  thresholdTable <- thresholdTable %>% map_df(rev)
  sftThreshold <- thresholdTable[which(thresholdTable$SFT > 0.9)[1],]$Threshold
  
  result <- list()
  result$suggestedThreshold <- suggestedThreshold
  result$sftThreshold <- sftThreshold
  result$thresholdTable <- thresholdTable
  return(result)
  
}
```

```{r}
plotThresholdMetrics <- function(seidrFile, thresholdTable, suggestedThreshold, sftThreshold, title="") {
  
  require(ggplot2)
  require(tibble)
  
  outPlot <- file.path(dirname(seidrFile), "threshold.png")
  plt <- as_tibble(reshape2::melt(thresholdTable, id.vars = c("Threshold")))
  
  p <- ggplot(plt, aes(x = Threshold, y = value)) +
    facet_wrap(~variable, scales = "free_y") + geom_line() +
    scale_x_reverse() + scale_y_continuous() +
    geom_vline(aes(xintercept = suggestedThreshold, colour=1)) +
    geom_vline(aes(xintercept = sftThreshold, colour=2)) +
    ggtitle(title) +
    theme(legend.position = "none")
  
  png(outPlot,
      units="px", width=1920, height=1080, res=300)
  print(p)
  dev.off()
  
  return(p)
}
```

```{r}
seidrApplyThreshold <- function(seidrFile, threshold) {
  folder <- dirname(seidrFile)
  outFile <- file.path(folder, "aggregatedThreshold.sf")
  
  # removing previous data
  if (file.exists(outFile)) {
    file.remove(outFile)
  }
  system(paste("seidr view", seidrFile, "-t", threshold, "-b -o", outFile), intern=TRUE)
  system(paste("seidr reheader", outFile), intern=TRUE)
}
```

This chunk can't be executed at the same time as knit generation, it causes conflict with seidr.
Instead we can run this in Rstudio and then save the object. In a step after we can load the object
and plot the results.

```{r, message=FALSE, warning=FALSE, eval=F}

tResults <- list() 

sapply(species, function(specie) {
  sapply(conditions, function(condition) {
    cat(paste("<h3>", specie, "-", condition, "</h3>"))

    seidrFile <- file.path(projectDir, specie, "networks", condition, "results/aggregated/aggregated.sf")
    
    tResults[[specie]][[condition]]$seidrFile <<- seidrFile
    
    # obtain recomended minimal threshold value
    minThreshold <- seidrDescribe(seidrFile)
    
    # obtain threshold values suggested, SFT based values and table
    thresholdValues <<- seidrThreshold(seidrFile, minThreshold)
    tResults[[specie]][[condition]]$thresholdValues <- thresholdValues
    
    # plot threshold metrics
    p <- plotThresholdMetrics(seidrFile, thresholdValues$thresholdTable, thresholdValues$suggestedThreshold, thresholdValues$sftThreshold, title = paste(specie, "-", condition))
    
    tResults[[specie]][[condition]]$plot <<- p
    
    # obtain thresholded seidr file
    seidrApplyThreshold(seidrFile, thresholdValues$sftThreshold)
    })
})

save(tResults, projectDir, species, conditions, file=file.path(projectDir, "thresholdResults.RData"))
```

```{r loadtResults, message=FALSE, warning=FALSE, results='asis'}
load(file=file.path(projectDir, "thresholdResults.RData"))

None <- sapply(species, function(specie) {
  sapply(conditions, function(condition) {
    cat(paste("<h3>", specie, "-", condition, "</h3>"))
    print(tResults[[specie]][[condition]]$plot)
    #tResults[[specie]][[condition]]$seidrFile
  })
})
```


# Network clustering
For each network, we will first some parameters, like folder location of the Seidr file, the markov-time parameter we found acceptable for clustering, the index edge file obtained from Seidr.

```{r getEdgeIndex, message=FALSE, warning=FALSE}
#load(file=file.path(projectDir, "thresholdResults.RData"))

None <- sapply(species, function(specie) {
  sapply(conditions, function(condition) {
    #cat(paste("<h3>", specie, "-", condition, "</h3>"))

    # get subfolder
    folder <- dirname(tResults[[specie]][[condition]]$seidrFile)
    seidrFile <- file.path(folder, "aggregatedThreshold.sf")
    
    # create the names of the files for Infomap
    edgeIndexFile <- file.path(folder, "indexEdgeList.txt")
    edgeFile <- file.path(folder, "edgeList.txt")
    treeFile <- file.path(folder,"indexEdgeList.tree")
    
    # get the index for the irp_score in Seidr view (it might different depending on how many methods are used)
    headResult <- system(paste("seidr view", seidrFile, "-c -d $'\t' ", "| head -n 1"), intern=TRUE)
    headResult <- unlist(strsplit(headResult, "\t"))
    algoIndex <- grep("irp_score", headResult)
    
    # 
    system(paste0("seidr view ", seidrFile, "  -d $'\t' | cut -f 1,2,",algoIndex, " > ",edgeFile), intern=TRUE)
    system(paste0("seidr view ", seidrFile, " -N -d $'\t' | cut -f 1,2,",algoIndex," >",edgeIndexFile), intern=TRUE)
    
    tResults[[specie]][[condition]]$edgeIndexFile <<- edgeIndexFile
    tResults[[specie]][[condition]]$edgeFile <<- edgeFile
    tResults[[specie]][[condition]]$treeFile <<- treeFile
    #tResults[[specie]][[condition]]$seidrFile
  })
})
```

```{r infomapCall, eval=F}
runInfomap <- function(seidrFile, edgeIndexFile, treeFile, markovTime=1, folder) {
  
  
  system(paste("Infomap", edgeIndexFile,"-z --markov-time", markovTime, folder))
  infomapRes <- system(paste("seidr resolve -s", seidrFile, treeFile), intern=TRUE)
  infomapTable <-data.frame(do.call(rbind, strsplit(infomapRes, "\t")))
  infomapTable <- prepareData(infomapTable)
  infomapTable$gene <- gsub("\r","",infomapTable$gene)
  infomapTable$Level1 <- infomapTable$P1
  infomapTable$Level2 <- paste0(infomapTable$Level1,":",infomapTable$P2)
  infomapTable$Level2 <- ifelse(infomapTable$Level2 %like% "NA", NA, infomapTable$Level2)

  if("P3" %in% colnames(infomapTable)) {
    infomapTable$Level3 <- paste0(infomapTable$Level1,":",infomapTable$P2, ":",infomapTable$P3)
    infomapTable$Level3 <- ifelse(infomapTable$Level3 %like% "NA", NA, infomapTable$Level3)
  }
  return(infomapTable)
}
```


```{r}

infomapResults <- list() 

None <- sapply(species, function(specie) {
  sapply(conditions, function(condition) {
    print(paste(specie, "-", condition))
    folder <- dirname(tResults[[specie]][[condition]]$seidrFile)
    seidrFile <- file.path(folder, "aggregatedThreshold.sf")
    edgeIndexFile <- tResults[[specie]][[condition]]$edgeIndexFile
    treeFile <- tResults[[specie]][[condition]]$treeFile

    infomapTable <- runInfomap(seidrFile, edgeIndexFile, treeFile, markovTime=1, folder) 
    
    infomapResults[[specie]][[condition]]$infomapTable <<- infomapTable
    infomapResults[[specie]][[condition]]$level1percent <<- clusterQA(infomapTable, level='Level1')
    infomapResults[[specie]][[condition]]$level2percent <<- clusterQA(infomapTable, level='Level2')
    
    if("P3" %in% colnames(infomapTable)) {
      infomapResults[[specie]][[condition]]$level3percent <<- clusterQA(infomapTable, level='Level3')
    }
    
  })
})

```

```{r}
None <- sapply(species, function(specie) {
  sapply(conditions, function(condition) {
    print(paste(specie, "-", condition))
    print(paste("Level1 %", infomapResults[[specie]][[condition]]$level1percent, "-", "Level2 %", infomapResults[[specie]][[condition]]$level2percent))
  })
})
```
The goal is to obtain ~50-60% of the genes in the top 20 clusters. Some of the candidates don't match our criteria. We need to focus individually in them and play with the markov-time parameter to match our criteria.

The candidates to recluster are:

"cenococcum_geophilum - control"
"cenococcum_geophilum - combined"
"cortinarius_glaucopus - control"
"hyaloscypha_variabilis - control"
"hyaloscypha_variabilis - combined"
"hyaloscypha_bicolor - control"
"piloderma_croceum - control"

Valid candidates
"cortinarius_glaucopus - combined" -> level2
"hyaloscypha_bicolor - combined" -> level1  
"piloderma_croceum - combined" -> leve2


## Recluster
```{r}
reCluster <- function(specie, condition, markovTime){
    folder <- dirname(tResults[[specie]][[condition]]$seidrFile)
    seidrFile <- file.path(folder, "aggregatedThreshold.sf")
    edgeIndexFile <- tResults[[specie]][[condition]]$edgeIndexFile
    treeFile <- tResults[[specie]][[condition]]$treeFile

    infomapTable <- runInfomap(seidrFile, edgeIndexFile, treeFile, markovTime, folder) 
    
    results <- list()
    results$infomapTable <- infomapTable
    results$level1percent <- clusterQA(infomapTable, level='Level1')
    results$level2percent <- clusterQA(infomapTable, level='Level2')
    print(paste("Level1 %", results$level1percent, "-", "Level2 %", results$level2percent))
    if("P3" %in% colnames(infomapTable)) {
      results$level3percent <- clusterQA(infomapTable, level='Level3')
      print(paste("Level1 %", results$level1percent, "-", "Level2 %", results$level2percent, "-", "Level3 %", results$level3percent))
    }
    
    return(results)
  }
```

For each candidate we will try to find a better markov-time to match our criteria

### "cenococcum_geophilum - control"
```{r, eval=FALSE}
  infomapPartialResults <- reCluster("cenococcum_geophilum", "control", 1.5)
  infomapResults$cenococcum_geophilum$control <- infomapPartialResults
```

"cenococcum_geophilum - combined"
```{r, eval=FALSE}
  infomapPartialResults <- reCluster("cenococcum_geophilum", "combined", 1.164)
  infomapResults$cenococcum_geophilum$combined <- infomapPartialResults
```


"cortinarius_glaucopus - control"
No better parameters found, we keep markov-time = 1 and take level 3 clusters
```{r message=FALSE, warning=FALSE, eval=FALSE}
  infomapPartialResults <- reCluster("cortinarius_glaucopus", "control", 1)
  infomapResults$cortinarius_glaucopus$control <- infomapPartialResults
```

"hyaloscypha_variabilis - control"
```{r message=FALSE, warning=FALSE, eval=FALSE}
  infomapPartialResults <- reCluster("hyaloscypha_variabilis", "control", 1.3)
  infomapResults$hyaloscypha_variabilis$control <- infomapPartialResults
```
"hyaloscypha_variabilis - combined"
We keep level2
```{r message=FALSE, warning=FALSE, eval=FALSE}
  infomapPartialResults <- reCluster("hyaloscypha_variabilis", "combined", 1.2)
  infomapResults$hyaloscypha_variabilis$combined <- infomapPartialResults
```

"hyaloscypha_bicolor - control"
We keep level2
```{r message=FALSE, warning=FALSE, eval=FALSE}
  infomapPartialResults <- reCluster("hyaloscypha_bicolor", "control", 1.3)
  infomapResults$hyaloscypha_bicolor$control <- infomapPartialResults
```

"piloderma_croceum - control"
We keep level3
```{r message=FALSE, warning=FALSE, eval=FALSE}
  infomapPartialResults <- reCluster("piloderma_croceum", "control", 1.2)
  infomapResults$piloderma_croceum$control <- infomapPartialResults
```

```{r eval=FALSE, include=FALSE}
save(infomapResults, projectDir, species, conditions, file=file.path(projectDir, "infomapResults.RData"))
```

```{r eval=T, include=FALSE}
load(file=file.path(projectDir, "infomapResults.RData"))
```


We can check again if all networks have at least one level matching the criteria

```{r}
None <- sapply(species, function(specie) {
  sapply(conditions, function(condition) {
    print(paste(specie, "-", condition))
    if("level3percent" %in% names(infomapResults[[specie]][[condition]])){
      print(paste("Level1 %", infomapResults[[specie]][[condition]]$level1percent, "-", 
                  "Level2 %", infomapResults[[specie]][[condition]]$level2percent,"-", 
                  "Level3 %", infomapResults[[specie]][[condition]]$level3percent))
    } else {
      print(paste("Level1 %", infomapResults[[specie]][[condition]]$level1percent, "-", "Level2 %", infomapResults[[specie]][[condition]]$level2percent))
    }
    
  })
})
```

## Choose level
We upgrade the result list with the chosen level
```{r}
infomapResults$cenococcum_geophilum$control$clusterLevel <- "Level1"
infomapResults$cenococcum_geophilum$combined$clusterLevel <- "Level1"

infomapResults$cortinarius_glaucopus$control$clusterLevel <- "Level3"
infomapResults$cortinarius_glaucopus$combined$clusterLevel <- "Level2"

infomapResults$hyaloscypha_variabilis$control$clusterLevel <- "Level1"
infomapResults$hyaloscypha_variabilis$combined$clusterLevel <- "Level1"

infomapResults$hyaloscypha_bicolor$control$clusterLevel <- "Level2"
infomapResults$hyaloscypha_bicolor$combined$clusterLevel <- "Level1"

infomapResults$piloderma_croceum$control$clusterLevel <- "Level3"
infomapResults$piloderma_croceum$combined$clusterLevel <- "Level2"
```

## Obtain clusters
Loop again to obtain the number of cluster by min size.
Initially we put a min size of 20 genes per cluster
```{r}
minSize=20

None <- sapply(species, function(specie) {
  sapply(conditions, function(condition) {
    print(paste(specie, "-", condition))
    
    selectedClusters <- getClusterByMinSize(infomapResults[[specie]][[condition]]$infomapTable, 
                                            level= infomapResults[[specie]][[condition]]$clusterLevel,
                                            min=minSize)
    
    infomapResults[[specie]][[condition]]$selectedClusters <<-selectedClusters

     print(selectedClusters)
  })
})
```

With a minimal value of 20 genes per cluster we get more than 17 clusters per network, we can proceed to extract them and export the results.
In order to get a manageable amount of clusters, we are going to limit the results to 30 clusters 

```{r}
maxClusters <- 30

None <- sapply(species, function(specie) {
  sapply(conditions, function(condition) {
    print(paste(specie, "-", condition))
    
    folder <- dirname(tResults[[specie]][[condition]]$seidrFile)
    
    theClusters <- getClusters(infomapResults[[specie]][[condition]]$infomapTable, 
                               level= infomapResults[[specie]][[condition]]$clusterLevel, 
                               numberOfClusters= ifelse(infomapResults[[specie]][[condition]]$selectedClusters < maxClusters,
                                                        infomapResults[[specie]][[condition]]$selectedClusters, maxClusters))
    
    # level 3 present some NA due to not having a gene in that level, they need to be removed
    theClusters <- lapply(theClusters, function(x){x<-x[!is.na(x)]})
    
    infomapResults[[specie]][[condition]]$clusters <<- theClusters
    save4Cytoscape(theClusters, file=file.path(folder,"infomapClusters.tsv"))

    
  })
})


```

ONLY level1 clustering, for pahte visualization
```{r}
minSize=20
maxClusters <- 30

None <- sapply(species, function(specie) {
  sapply(conditions, function(condition) {
    print(paste(specie, "-", condition))
    
    selectedClusters <- getClusterByMinSize(infomapResults[[specie]][[condition]]$infomapTable, 
                                            level= 'Level1',
                                            min=minSize)
    
    infomapResults[[specie]][[condition]]$selectedClusters <<-selectedClusters

    folder <- dirname(tResults[[specie]][[condition]]$seidrFile)
    
    theClusters <- getClusters(infomapResults[[specie]][[condition]]$infomapTable, 
                               level= infomapResults[[specie]][[condition]]$clusterLevel, 
                               numberOfClusters= ifelse(infomapResults[[specie]][[condition]]$selectedClusters < maxClusters,
                                                        infomapResults[[specie]][[condition]]$selectedClusters, maxClusters))
    
    # level 3 present some NA due to not having a gene in that level, they need to be removed
    theClusters <- lapply(theClusters, function(x){x<-x[!is.na(x)]})
    
    infomapResults[[specie]][[condition]]$clusters <<- theClusters
    save4Cytoscape(theClusters, file=file.path(folder,"infomapClustersLevel1.tsv"))

    
  })
})
```


# Profile splitting

Adding the paths for the data
```{r}

infomapResults$cenococcum_geophilum$control$data <- file.path(projectDir, "cenococcum_geophilum",  "DE", "controlVsdData.RData")
infomapResults$cenococcum_geophilum$combined$data <- file.path(projectDir, "cenococcum_geophilum",  "DE", "combinedVsdData.RData")

infomapResults$cortinarius_glaucopus$control$data <- file.path(projectDir, "cortinarius_glaucopus",  "DE", "controlVsdData.RData")
infomapResults$cortinarius_glaucopus$combined$data <- file.path(projectDir, "cortinarius_glaucopus",  "DE", "combinedVsdData.RData")

infomapResults$hyaloscypha_variabilis$control$data <- file.path(projectDir, "hyaloscypha_variabilis", "DE", "controlVsdData.RData")
infomapResults$hyaloscypha_variabilis$combined$data <- file.path(projectDir, "hyaloscypha_variabilis",  "DE", "combinedVsdData.RData")

infomapResults$hyaloscypha_bicolor$control$data <- file.path(projectDir, "hyaloscypha_bicolor", "DE","controlVsdData.RData")
infomapResults$hyaloscypha_bicolor$combined$data <- file.path(projectDir, "hyaloscypha_bicolor", "DE", "combinedVsdData.RData")

infomapResults$piloderma_croceum$control$data <- file.path(projectDir, "piloderma_croceum", "DE", "controlVsdData.RData")
infomapResults$piloderma_croceum$combined$data <- file.path(projectDir, "piloderma_croceum", "DE","combinedVsdData.RData")
```


```{r splitClusters}
splitClusters <- function(infomapClusters, data) {
  
  result <- list()
  
  None <- sapply(names(infomapClusters), function(cluster){
    
    posCluster <- cluster
    negCluster <- paste0(cluster,"i")
    
    res <- getEigenGenes(data, infomapClusters[[cluster]])
    
    result[[posCluster]] <<- res$positive
    result[[negCluster]] <<- res$negative

  })
  
  
  return(result)

}

```

```{r }
None <- sapply(species, function(specie) {
  sapply(conditions, function(condition) {
    print(paste(specie, "-", condition))
    
    folder <- dirname(tResults[[specie]][[condition]]$seidrFile)
    
    load(infomapResults[[specie]][[condition]]$data)
    thisData <- NA
    if (condition == "control")
      thisData <- controlData
    if (condition == "combined")
      thisData <- combinedData
    
    splitRes <- splitClusters(infomapResults[[specie]][[condition]]$clusters, thisData)

    res <- melt(splitRes)

    names(res) <- c("gene","cluster")
    
    infomapResults[[specie]][[condition]]$splitClusters <<- res
    
    write.table(res, file=file.path(folder,"splitedClusters.tsv"), sep='\t', row.names = F, quote = F)

  })
})
```

```{r}
save(infomapResults, projectDir, species, conditions, file=file.path(projectDir, "infomapFinalResults.RData"))
```


```{r}

```

